### 2016-11-15 paper.tex
5.3 space saving

\documentclass[10pt,preprint]{sigplanconf}

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{stmaryrd}
\usepackage{color}
\usepackage{graphics}
\usepackage{fancyvrb}
\usepackage{subfigure}
\usepackage{multirow}
\usepackage{minted}
\usepackage{breqn}
\usepackage{enumitem}
%\usepackage{subcaption}
\usepackage{siunitx} % For pretty-printing numeric values and SI units
                     % of measure. e.g., the tabular column type S is
                     % used to print nice-looking tables of numbers.
\sisetup{ % defaults
  group-separator={,},
  group-minimum-digits={3},
  output-decimal-marker={.},
  table-format = 6
}

\usepackage{tikz}
\usepackage{xypic}
\usepackage{natbib}

\usetikzlibrary{calc}

%% http://tex.stackexchange.com/questions/55068/is-there-a-tikz-equivalent-to-the-pstricks-ncbar-command
\tikzset{
    ncbar angle/.initial=90,
    ncbar/.style={
        to path=(\tikztostart)
        -- ($(\tikztostart)!#1!\pgfkeysvalueof{/tikz/ncbar angle}:(\tikztotarget)$)
        -- ($(\tikztotarget)!($(\tikztostart)!#1!\pgfkeysvalueof{/tikz/ncbar angle}:(\tikztotarget)$)!\pgfkeysvalueof{/tikz/ncbar angle}:(\tikztostart)$)
        -- (\tikztotarget)
    },
    ncbar/.default=0.5cm,
}

\tikzset{round left paren/.style={ncbar=0.4cm,out=110,in=-110}}
\tikzset{round right paren/.style={ncbar=0.4cm,out=70,in=-70}}

\fvset{
  linenos=true,
  fontsize=\footnotesize,
  breaklines=true,
 breakafter=w),
  xleftmargin=\parindent
}

\usemintedstyle{vs}

\definecolor{darkblue}{rgb}{0.0,0.0,0.5}
\definecolor{darkgreen}{rgb}{0.0,0.4,0.0}
\definecolor{darkdarkgreen}{rgb}{0.0,0.35,0.0}

\usepackage{hyperref}
\hypersetup{
    unicode=false,          % non-Latin characters in Acrobat's bookmarks
    pdftoolbar=true,        % show Acrobat toolbar?
    pdfmenubar=true,        % show Acrobat menu?
    pdffitwindow=false,      % page fit to window when opened
    pdftitle={},    % title
    pdfauthor={}
    pdfsubject={},   % subject of the document
    pdfnewwindow=true,      % links in new window
    pdfkeywords={keywords}, % list of keywords
    colorlinks=true,       % false: boxed links; true: colored links
    linkcolor=darkblue,          % color of internal links
    citecolor=darkblue,        % color of links to bibliography
    filecolor=green,      % color of file links
    urlcolor=blue,          % color of external links
}

\CustomVerbatimEnvironment{SpecVerbatim}{Verbatim}{fontsize=\footnotesize,xleftmargin=0.5cm,
xrightmargin=0.2cm,commandchars=\\\{\},baselinestretch=0.98,numbersep=0.9em}
\CustomVerbatimEnvironment{ExmVerbatim}{Verbatim}{fontsize=\footnotesize,xleftmargin=0.5cm,
xrightmargin=0.2cm,baselinestretch=0.98,numbers=left,numbersep=0.9em,commandchars=\\\{\}}
\CustomVerbatimEnvironment{IVerbatim}{Verbatim}{fontsize=\relsize{-1},xleftmargin=0.5cm,
xrightmargin=0.2cm,commandchars=\\\{\},baselinestretch=0.98,numbersep=0.9em}

\definecolor{darkgreen}{rgb}{0.0,0.5,0.0}
\definecolor{darkpurple}{rgb}{0.6,0.0,0.6}
\definecolor{orange}{rgb}{0.8,0.4,0.0}
\definecolor{darkorange}{rgb}{0.5,0.2,0.0}
\definecolor{marco}{rgb}{0.0,0.3,0.5}
\definecolor{gray}{rgb}{0.2,0.2,0.2}

\newcommand\mdef{\stackrel{\mathclap{\normalfont\mbox{def}}}{=}}

\newcommand{\bn}{\mathbb{N}}

\newcommand{\todo}[1]{\textcolor{blue}{Dom: #1}}
\newcommand{\dnote}[1]{\textcolor{darkpurple}{Dom: #1}}
\newcommand{\mnote}[1]{\textcolor{darkgreen}{Mistral: #1}}
\newcommand{\anote}[1]{\textcolor{red}{Andy: #1}}

\newcounter{block}

\newtheorem{lemma}[block]{Lemma}
\newtheorem{proposition}[block]{Proposition}

\theoremstyle{definition}

\newtheorem{theorem}[block]{Theorem}
\newtheorem{remark}[block]{Remark}
\newtheorem{example}[block]{Example}
\newtheorem{definition}[block]{Definition}

% Writing macros
\newcommand{\ie}{\emph{i.e.}}
\newcommand{\eg}{\emph{e.g.}}

\newcommand{\dimId}{\texttt{dim}}

% Semantics related
\newcommand{\interp}[1]{\llbracket{#1}\rrbracket}

% Syntax macros
\newcommand{\nonterm}[1]{\textit{#1}}
\newcommand{\term}[1]{\texttt{#1}}

\newcommand{\stenRefl}[1]{\term{pointed} \, (\term{dim=}#1)}
\newcommand{\stenFwd}[3]{\term{forward} \, (\term{depth=}#1,
  \term{dim=}#2{#3})}
\newcommand{\stenBwd}[3]{\term{backward} \, (\term{depth=}#1,
  \term{dim=}#2{#3})}
\newcommand{\stenCen}[3]{\term{centered} \, (\term{depth=}#1,
  \term{dim=}#2{#3})}
\newcommand{\irrefl}{\texttt{nonpointed}}

\newcommand{\stenReflS}[1]{\term{pointed} \, (\term{dim=}#1)}
\newcommand{\stenFwdS}[2]{\term{fwd} \, (\term{depth=}#1,
  \term{dim=}#2)}
\newcommand{\stenBwdS}[2]{\term{bwd} \, (\term{depth=}#1,
  \term{dim=}#2)}
\newcommand{\stenCenS}[2]{\term{cen} \, (\term{depth=}#1,
  \term{dim=}#2)}
\newcommand{\irreflS}{\texttt{np}}

\newcommand{\stenFwdSR}[3]{\term{fwd} (\term{depth=}#1,
  \term{dim=}#2, #3)}
\newcommand{\stenBwdSR}[3]{\term{bwd} (\term{depth=}#1,
  \term{dim=}#2, #3)}
\newcommand{\stenCenSR}[3]{\term{cen} (\term{depth=}#1,
  \term{dim=}#2, #3)}
\newcommand{\stenReflSR}[1]{\term{point} (\term{dim=}#1)}

% SYNTAX OPERATIONS AND PREDICATES
%\newcommand{\neigh}{\textsf{neigh}}
\newcommand{\arrayTy}{\textsf{array}}
\newcommand{\rhsExp}{\textsf{rhsExp}}
\newcommand{\var}{\textsf{var}}

%% VECTOR NOTATIONS
\newcommand{\vect}[1]{\textbf{#1}}

\newcommand{\vtwohs}[2]{\setlength{\arraycolsep}{0em}
\left[\begin{array}{cc}#1 \, & \, #2\end{array}\right]\!}

\newcommand{\vtwoh}[2]{\setlength{\arraycolsep}{0em}
\left[\begin{array}{cc}#1 \; & \; #2\end{array}\right]}

\newcommand{\vthreeh}[3]{\setlength{\arraycolsep}{0em}
\left[\begin{array}{ccc}#1 \; & \; #2 \; & \; #3\end{array}\right]}
\newcommand{\vtwo}[2]{\setlength{\arraycolsep}{0em}
\left[\begin{array}{l}#1\\#2\end{array}\right]}
\newcommand{\vthree}[3]{\setlength{\arraycolsep}{0em}
\left[\begin{array}{l}$#1$\\$#2$\\$#3$\end{array}\right]}
\newcommand{\stwo}[4]
%{\vtwo{#1}{#2}\!\vtwo{#3}{#4}}
{\setlength{\arraycolsep}{0.1em}
\left[\begin{array}{rr}$#1$ & $#3$\\$#2$ & $#4$\end{array}\right]}

\newcommand{\singleEntry}[2]{\textbf{J}_{#2}^{#1}}
\newcommand{\zeroEntry}[2]{\textbf{K}_{#2}^{#1}}

%% OPERATIONS ON SPANS and VECTORS
\newcommand{\containedin}{\sqsubseteq}

%% MODEL
\newcommand{\effdims}[2]{\mathit{constr}(#1)_{#2}}

\newcommand{\trule}[1]{{\footnotesize{(\text{#1})}}}
%\include{results}

%\newcommand{\cons}[3]{\mathbf{C}_{#3}(#1, #2)}
\newcommand{\consName}{\textbf{agree\textprime}}
\newcommand{\cons}[2]{#1 \,\, \consName{} \,\, #2}
\newcommand{\consAName}{\textbf{agree}}
\newcommand{\consA}[2]{#1 \,\, \consAName{} \,\, #2}
\newcommand{\consSub}[3]{#2 \,\, \consAName{}_{#1} \,\,#3}

\title{Verifying Spatial Properties of Stencil Computations}
%\titlebanner{DRAFT -- Do not distribute}
\authorinfo{}{}{}
%\preprintfooter{DO NOT DISTRIBUTE}

\begin{document}
\maketitle

\bibliographystyle{abbrvnat}

\begin{abstract}
Stencil computations are a common programming pattern in numerical
  code and typically form the core of graphics and scientific
  computing applications. However, they are often complex and
  unwieldy, involving substantial fine-grained manipulation of array
  indices across multiple arrays and dimensions. This requires careful
  testing, which is almost always done manually and then discarded
  during development. In this paper, we propose an alternate approach:
  a specification language and automated tool for verifying the
  spatial behaviour of stencils. The specification language is based
  on the hypothesis that stencil computations read from arrays with a
  regular, fixed pattern which can be captured abstractly by a simple
  set of combinators in our specification language. These
  specifications can be embedded as annotations in source code,
  against which code can be checked for conformance. For legacy code,
  spatial specifications can be inferred, providing documentation and
  aiding future software maintenance. We evaluate our language against
  a corpus of numerical Fortran code, and show that the vast majority
  of stencil computations indeed have a simple, regular, static shape
  validating the design of our language.  We introduce the
  language, checking and inference procedures, and its evaluation.
\end{abstract}

%\keywords{program verification, specification, stencils}

\section{Introduction}\label{sec:intro}

\emph{Stencils} are a ubiquitous programming pattern, common in
scientific and numerical computing applications. Informally, a stencil
computation computes an array whose value at each index $i$ is
calculated from a \emph{neighbourhood} of values around $i$ in some
input array(s), \eg{}, the Game of Life, convolutions in image
processing, approximations to differential equations. For example, the
following computes the one-dimensional discrete Laplace transform (an
approximation to a derivative) in Fortran:
%
\begin{minted}{fortran}
do i = 1, (n-1)
      b(i) = a(i-1) - 2*a(i) + a(i+1)
end do
\end{minted}
%
Values \texttt{b(i)} are calculated from a neighbourhood of elements
around \texttt{i} in the input array \texttt{a}.  In this example, the
access pattern is simple and easily understood. More complex
stencil computations are more prone to errors from simple lexical
mistakes. For example, the following is a small snippet from a
Navier-Stokes fluid simulator (based on~\citet{griebel1997numerical})
 in which two arrays are read with
different data access patterns, across two dimensions.
%%
\begin{minted}[firstnumber=20,xleftmargin=2em]{fortran}
du2dx = ((u(i,j)+u(i+1,j))*(u(i,j)+u(i+1,j))+   &
  gamma*abs(u(i,j)+u(i+1,j))*(u(i,j)-u(i+1,j))- &
  (u(i-1,j)+u(i,j))*(u(i-1,j)+u(i,j))-          &
  gamma*abs(u(i-1,j)+u(i,j))*(u(i-1,j)-u(i,j))) &
  /(4.0*delx)

duvdy = ((v(i,j)+v(i+1,j))*(u(i,j)+u(i,j+1))+   &
  gamma*abs(v(i,j)+v(i+1,j))*(u(i,j)-u(i,j+1))- &
  (v(i,j-1)+v(i+1,j-1))*(u(i,j-1)+u(i,j))-      &
  gamma*abs(v(i,j-1)+v(i+1,j-1))*(u(i,j-1)-     &
  u(i,j))) / (4.0*dely)

laplu = (u(i+1,j)-2.0*u(i,j)+u(i-1,j))/delx/delx+ &
  (u(i,j+1)-2.0*u(i,j)+u(i,j-1))/dely/dely

f(i,j) = u(i,j)+del_t*(laplu/Re-du2dx-duvdy)
\end{minted}
%%
The access pattern is much harder to understand than the
one-dimensional Laplace.  The miasma of indexing expressions of the
form \texttt{var(i$\pm$a, j$\pm$b)} is not only hard to read, but is
prone to simple textual input mistakes, \eg{}, swapping \texttt{-} and
\texttt{+}, missing an indexing term, or transforming the wrong
variable \eg{} \texttt{(i+1,j)} instead of \texttt{(i,j+1)}.

In practice, the typical development procedure for complex
stencil computations involves some ad hoc testing to ensure that no
transcription mistakes have been made. This may involve visual
inspection of the results, based on a rendered image. Typically such
testing is discarded once the code is seen to be correct.

This is not the only information that is often discarded. The
``shape'' of the indexing pattern is usually the result of choices made in the
numerical-analysis procedure to convert some
continuous equations into a discrete approximation. Rarely are these
decisions captured in the source code, yet the shape of access is
usually uniform and is has a perspicuous and concise description in
numerical analysis literature \eg{}, ``\emph{centered in space, to a
  depth of 1}'' referring to indexing terms \texttt{a(i)},
\texttt{a(i-1)} and \texttt{a(i+1)}.

To support the development of correct stencil computations, we propose
a simple, abstract specification language for the data access pattern
of stencils. The shape of the stencil computation in the
Laplace example is specifed in our language as:
%
\begin{minted}[linenos=false]{fortran}
!=  stencil centered(depth=1, dim=1) :: a
\end{minted}
%
This describes that \texttt{a} is accessed with a symmetrical pattern
(``centered'') to a depth of one in each direction in its first
dimension.  The shape of the Navier-Stokes example
is specified as:
%
\begin{minted}[linenos=false,xleftmargin=0em,xrightmargin=0.8em,breakindent=1.9em]{fortran}
!= stencil centered(depth=1,dim=1)*pointed(dim=2)     + centered(depth=1,dim=2)*pointed(dim=1) :: u

!= stencil forward(depth=1,dim=1)                          * backward(depth=1,dim=2) :: v
\end{minted}
%
The specification requires that, over the whole fragment, \texttt{u}
is accessed with a centered pattern to depth of 1 in both dimensions
(this is known as the \emph{five-point stencil}) and \texttt{v} is
accessed in a neighbourhood bounded forwards to depth of $1$ in the
first dimension and backward to a depth of $1$ in the second
dimension. Note the relatively small size of the specification
compared to the source code, and its high-level of abstraction.

We provide a verification tool which can check the spatial correctness
of stencil computations against a specification. Furthermore, the tool
can automatically infer specifications and insert these as comments
into the source code. The main features and benefits of our approach
are as follows:
%
\begin{itemize}
%\item The main use case is that a programmer writes a specification
%  first (perhaps informed by their numerical analysis), then writes
%  the corresponding source code, and lastly has the tool check its
%  conformance to the specifications. This specify-and-check approach
%  removes the usually ad hoc testing effort, and provides the
%  opportunity for easy re-verification in the event of any refactoring
%  of software maintenance activities.
\item The primary use case is that a programmer writes a specification first,
  then writes the corresponding source code. Our tool then checks conformance between
  the two. This specify-and-check approach reduces testing effort and
  future-proofs against bugs introduced during refactoring and maintanence.
%
%\item The specification provides a concise description which can be
%  preserved in the source code to provide documentation to future
%  developers, contrasting with ad hoc tests which are usually
%  discarded or specialised to the development environment. Separate from the
%  verification procedure, the infer-and-synthesise mode of the tool
%  provides a way to automatically generate additional useful
%  documentation for future users.

% Emphasis here is on documentation generation, so it makes little sense to
% compare this aspect of CamFort to ad hoc testing. Apples and oranges.
\item A specification concisely captures the access pattern of a
  stencil and turns it into documentation. The inference procedure
  efficiently produces specifications with no programmer intervention,
  automatically inserting specifications at appropriate places in the
  source code.
%
%\item Specifications are integrated into the code as comments for easy
%  integration with existing tools, as well as documentation systems (\eg{}, we
%  support Doxygen and FORD syntax).
%
\item Our specification format is deliberately abstract, with a small number of
  combinators that \emph{do not involve any indexing expressions}, \eg{}
  \texttt{a(i+1,j-1)}. This contrasts with other specification approaches. If a
  user specifies, say, the Naiver-Stokes stencil above, in a deductive
  verification language such as ACSL, they must write a specification involving
  textually dense array-indexing expressions. Therefore, any low-level mistakes
  that could be made whilst programming complex indexing code could also be
  made when writing its specification. Our specifications are more lightweight
  with the aim of aiding adoption by scientists and preventing indexing errors.
\end{itemize}

Our verification tool does not target the
class of bugs that can be detected automatically (\emph{push-button
  verification}). Instead, stencil computation bugs must be
identified relative to a specification of the intended access pattern.

In this paper, we make the following contributions:
%
\begin{itemize}
\item We introduce a specification language for stencils
  that captures common forms of data access patterns
  (\S~\ref{sec:lang});

\item We detail checking and inference algorithms for our
  specifications (\S~\ref{sec:analysis}), derived from a
  denotational model for the specificaiton language
  (\S~\ref{sec:semantics});

\item we provide an implementation of our approach as an extension to
  CamFort, an open-source program analysis tool for Fortran;

\item We report on a quantitative study of stencil computations on a
  corpus of numerical Fortran program
  (\S~\ref{sec:evaluation}), totalling one million lines of code. Our tool
  identifies and infers specifications for 35,000 stencil computations in the
  corpus. Approximately 10\% of the stencils we found are non-trivial,
  corresponding to code which is a possible source of errors. This validates
  our hypothesis that the majority of stencil computations have a regular
  shape, and validates the design of our language in its capability to capture
  many core patterns.

\item We give a detailed verification case study of two
  stencil computations commonly used in scientifc computing,
  simulating programming errors. Our approach detects all
  possible indexing errors for these computations.
\end{itemize}
%

\newcommand{\domainVal}{\mathbb{Z}_\infty}
\section{Stencil specifications}

Our specification language is based on the hypothesis
that most forms of array access in numerical code have
a fixed, statically-determined access pattern. For example, the
\emph{five-point stencil} on a two-dimensional array reads from array
indices $(i, j)$, $(i-1, j)$, $(i+1, j)$, $(i, j-1)$, and $(i, j+1)$
for all $i, j$ within the inner boundary of the array (to avoid
out-of-bounds access at the edges). We revisit this hypothesis
in \S~\ref{sec:evaluation} with the inference of
such regular stencil patterns on our corpus of numerical programs.

We give various definitions used in the paper then
\S~\ref{sec:syntax} outlines the syntax of stencil specifications
with an informal introduction. \S~\ref{sec:semantics}
gives a semantic model via a set-based interpretation which makes
concrete the meaning of specifications. \S~\ref{sec:eqs} then provides
a syntactic equational theory for specifications which is
sound with respect to the model.
%sub-specification for which the
%equality and sub-specification relation are proven sound.

\paragraph{Notation}
%\label{sec:notation}

\renewcommand*{\arraystretch}{0.8}
For the target language, $e$ ranges over expressions (may be impure) and $v$
over its (imperative) variables. %By
%``variable'', we mean in the imperative sense (named binders to memory cells).

\begin{definition}[Induction variable]
  An integer variable is an \emph{induction variable} if it is the control
  variable of a \texttt{for} loop incremented by $1$ per iteration. A variable
  is interpreted as an induction variable only within the scope of the loop
  body. Induction varaibles range over $i, j, k$ throughout.
%TODO: consider mentioning derived, non-based induction variables
% as we did before.
\end{definition}

\begin{definition}[Array subscripts and indices]
  An \emph{array subscript}, denoted $a(\bar{e})$, is an expression
  which reads from an array $a$ at an \emph{index} specified by a
  comma-separated sequence of integer expressions denoted $\bar{e}$ or
  in expanded form as $(e_1, \ldots, e_n)$. An index $\bar{e}$
  is called \emph{relative} if each $e \in \bar{e}$ involves an induction
  variable.

  Absolute index is a constant integer expression relative to the enclosing
  loop.
\end{definition}

\begin{definition}[Neighbourhood index]
\label{def:neighbour}
  For an array subscript $a(\bar{e})$ an index $\bar{e}$ is called a
  \emph{neighbourhood index} if every $e \in \bar{e}$ is of the form
  $e \equiv i$ or $e \equiv i + c$ or $e \equiv i - c$ where $c$ is an
  integer constant. That is, a neighbourhood index is a relative index
  where each component is a constant translation of a base induction
  variable. (The relation $\equiv$ here identifies terms up-to
  commutativity of $+$ and the inverse relationship of $+$ and $-$
  \eg{}, $(-b) + i \equiv i - b$).
\end{definition}

\begin{definition}[Stencil computations]
\label{def:stencil}
  Let $a$ be an array of dimensionality $n$ and $\overline{b}$ a
  collection of arrays of arbitrary (possible differing)
  dimensionalities.
  A \emph{stencil computation} comprises an iteration by a set $I$ of at
  most $n$ induction variables
  over a subset of the index space of $a$. Each step of the interation
  determs the elements of $a$ by an
  assignment $a(\bar{e}) = e_r$ where each $e \in
  \bar{e}$ is either a neighbourhood or constant index.
  For all arrays $b \in \bar{b}$ and array subscripts $b(\bar{e}')$ in
  $e_r$ and that flow to $e_r$, then each $e' \in \bar{e}'$ is also
  either a neighbourhood or constant index.

  We refer to the assignment $a(\bar{e}) = e_r$ and all the associated
  statements with dataflow to $e_r$ as the \emph{stencil kernel}.
\end{definition}
%
% specifying the shape of indexing expressions for
%arrays $\bar{v}$ that are simultaenously permitted and expected within
%the code to which the specification is given.
%
%\marginpar{\dnote{Is it worth having this informal consistency here?}}
\begin{definition}[Consistency, informal]
Specifications are associated to a stencil kernel
and are declared for one or more array variables
$\bar{v}$ within that kernel.
  A stencil specification defines a \emph{region} associated to an
  array variable $a$ against which a stencil kernel should be
  \emph{consistent}. Informally, the region defined by a specification
  is an $n$-dimensional rectilinear shape in a discrete space crossing the origin with respect to the induction
  variable. Consistency is then a two way requirement.
  On the one hand, all array subscripts on $a$ contributing to the kernel
  should be consistent with the region; on the other, all parts of the
  region should have an index corresponding to that part.
\end{definition}

\begin{figure}[t]
\vspace{-0.9em}
\begin{align*}
\def\arraystretch{1}
\setlength{\arraycolsep}{0.2em}
\newcommand{\dimTy}{\mathbb{N}_{>0}}
\begin{array}{rl}
\nonterm{specification} ::= & \nonterm{regionDec} \mid \nonterm{specDec} \\
\nonterm{specDec} ::= & \term{stencil} \; \nonterm{spec} \;
                        \texttt{::} \; v \\
\nonterm{regionDec} ::= &  \texttt{region} \; \texttt{::} \; \nonterm{rvar} \; \texttt{=} \;
                         \nonterm{region}\\[0.4em]
%\nonterm{spec} ::= & \nonterm{spatial} \mid \nonterm{temporal}
%\\[1em]
\nonterm{spec} ::= & [\nonterm{mult},] \; [\nonterm{approx},] \; \nonterm{region} \\
\nonterm{mult} ::= & \term{readOnce} \\
\nonterm{approx} ::= & \term{atMost} \; \mid \; \term{atLeast} \\[0.1em]
\nonterm{region} ::= & \nonterm{rvar} \; \\
\multicolumn{2}{l}{\qquad\qquad \mid \; \stenRefl{\dimTy}} \\
\multicolumn{2}{l}{\qquad\qquad \mid \;\stenFwd{\mathbb{N}_{>0}}{\dimTy}{\;[, \texttt{nonpointed}]}} \\
\multicolumn{2}{l}{\qquad\qquad \mid \; \stenBwd{\mathbb{N}_{>0}}{\dimTy}{\;[, \texttt{nonpointed}]}} \\
\multicolumn{2}{l}{\qquad\qquad \mid \; \stenCen{\mathbb{N}_{>0}}{\dimTy}{\;[, \texttt{nonpointed}]}} \\
\multicolumn{2}{l}{\qquad\qquad \mid \; \nonterm{region} \, \term{+}
  \, \nonterm{region} \; \mid \; \nonterm{region} \; \term{*} \; \nonterm{region}} \\[0.25em]
%\multicolumn{2}{l}{\qquad\qquad \mid \; \nonterm{rvar}}
%\\[0.5em]
%\nonterm{temporal} ::= \; & \term{dependency} \; (v \; \{ , v \}) [, \texttt{mutual}]
%  \\[0.5em]
\nonterm{rvar} ::= \; & [\text{\term{a}-\term{z}$\,$\term{A}-\term{Z}$\,$\term{0}-\term{9}}]+\\[-1em]
\end{array}
\end{align*}
\caption{Specification syntax (EBNF grammar)}
\label{fig:syntax}
\vspace{-0.8em}
\end{figure}

\subsection{Specification syntax}
\label{sec:syntax}

Figure~\ref{fig:syntax} gives the syntax of stencil specifications,
which we introduce in stages below.  The top-level is given by the
\nonterm{specification} production which splits into either a
\nonterm{regionDec} (region declaration) or a \nonterm{specDec}
(specification declaration).

Regions comprise \emph{region constants} which are
combined via region operators \term{+} and \term{*}.

\paragraph{Region constants}

Region constants specify a finite contiguous region in a
single dimension relative to the origin and are either \term{pointed},
\term{forward}, \term{backward}, or \term{centered}.  Each region
constant has a dimension identifier $d$ given as a natural number
greater than 0. Each constant except \term{pointed} has a depth
parameter $n$ given as a natural number greater than 0; \term{pointed}
regions implicitly have a depth of $0$.

A \term{forward} region of depth $n$ specifies a continguous
region in dimension $d$ starting at the origin; this corresponds
to specifying neighbourhood indices in dimension $d$ ranging from $i$ to $i + n$
inclusively for some induction variable $i$.  Similarly, a
\term{backward} region of depth $n$ corresponds to contiguous indices
in dimension $d$ from $i$ to $i - n$, and \term{centered}
of depth $n$ from $i - n$ to $i + n$. A \term{pointed}
stencil specifies a neighbour index $i$. For example the
following shows four specifications with four consistent stencil
kernels reading from arrays \term{a}, \term{b}, \term{c} and \term{d} respectively:
%%
\begin{minted}{fortran}
!= stencil forward(depth=2, dim=1) :: a
e(i, 0) = a(i, 0) + a(i+1, 0) + a(i+2, 0)
!= stencil backward(depth=2, dim=1) :: b
e(i, j) = b(i, j) + b(i-1, j) + b(i-2, j)
!= stencil centered(depth=1, dim=2) :: c
e(i, j) = (c(j-1) + c(j) + c(j+1))/3.0
!= stencil pointed(dim=3) :: d
e(i, j) = d(0, 0, i)
\end{minted}
%%
Not every dimension needs to be specified. In the above, the
specifications on line $1$, $3$, and $7$ leaves some dimensions unspecified.

The \term{forward}, \term{backward}, and \term{centered} regions may
all have an additional attribute \term{nonpointed} which marks that
the region does not include the origin.  For example, the following is a
\term{nonpointed} \term{backward} stencil
%
\begin{minted}{fortran}
!= stencil backward(depth=2, dim=1, nonpointed) :: a
b(i) = a(i-1) + a(i-2)
\end{minted}
%
The \term{pointed} regions cannot have a \term{nonpointed} attribute.

\paragraph{Combining regions}

% r+s means that indices must be consistent with either r or s
% r*s means that indices must be consistent with both r and s at the
% same time

% specifications correspond to a set of schemes

% r+s means an index must be generalised by a scheme in either r or s
%       and every scheme in r must have an index which specialises it
%       and every scheme in s must have an index which specialises it

% r*s means an index must be generalised by a scheme in r and a scheme in s
%       and every pairs of schemes in r and s must have an index which
%           specialises both
%

The region operators \term{+} and \term{*} provide ways to combine
regions disjunctively and conjuctively respectively.

The conjuction of two regions $r \term{*} s$ means that any indices
in the specified code must be consistent with both $r$ and $s$
simultaenously. Dually, for every pair of components in the regions
$r$ and $s$ there must be an index consistent with both.
For example, the following code (defining a \emph{nine-point stencil})
has a specification given by the product of two \texttt{centered}
regions in each dimension:
%%
\begin{minted}[breakindent=2.9em]{fortran}
x = a(i, j)   + a(i-1, j)   + a(i+1, j)
y = a(i, j-1) + a(i-1, j-1) + a(i+1, j-1)
z = a(i, j+1) + a(i-1, j+1) + a(i+1, j+1)
!= stencil centered(depth=1, dim=1) * centered(depth=1, dim=2) :: a
b(i, j) = (x + y + z) / 9.0
\end{minted}
%%
This pattern is common in image convolution applications (for example
for edge detection). The specification ranges over the
values that flow to the array subscript on the left-hand side,
and so ranges over the intermediate assignments to \term{x},
\term{y}, and \term{z}. Each index in the code is consistent
with both specifications simultaneously, \eg{}, \texttt{a(i-1, j+1)}
is within the centered region in dimension $1$ and the centered region
in dimension $2$.

The product $r \term{*} s$ can also be thought of as a bounding box
over the two regions $r$ and $s$.

The disjunction of two regions $r \term{+} s$ means that any indices
in the spcified code must be consistent with either of $r$ or $s$.
Dually, for every component in the region $r$ there must be an index
within it and every component in the region $s$ must have an index
within it. For
example, the following gives the specification of a five-point
stencil which is the sum of two compound \texttt{pointed} and
\texttt{centered} regions in each dimension:
%
\begin{minted}{fortran}
!= stencil centered(depth=1, dim=1)*pointed(dim=2)    + centered(depth=1, dim=2)*pointed(dim=1) :: a
b(i,j) = -4*a(i,j) + a(i-1,j) + a(i+1,j) &
                   + a(i,j-1) + a(i,j+1)
\end{minted}
%%
Here the left-hand side of \texttt{+} says that when the second dimension
(induction variable $j$) is fixed at the origin, the first dimension
(induction variable $i$) accesses the immediate vicinity of the origin
(to depth of one). The right hand side of \texttt{+} is similar but the dimensions are reversed.
This reflects the symmetry under rotation of the five-point stencil.

As another example, the following
implements and specifies the \emph{Roberts cross}
edge-detection convolution~\cite{davis1975survey}:
%%
\begin{minted}[breakindent=5.5em]{fortran}
do j=0, jmax-1
  do i=0, imax-1
    x = a(i,j)-a(i+1,j+1); y = a(i+1,j)-a(i,j+1)
    != stencil forward(depth=1,dim=1) * forward(depth=1,dim=2) :: a
    b(i,j) = sqrt(x*x+y*y)
  end do
end do
\end{minted}

\paragraph{Region declarations and variables}

Region specifications can be assigned to region variables via
a region declaration, which can be used later in another
specification. For example, the last specification
 example above can be restated as:
%%
\begin{minted}{fortran}
!= region :: r1 = forward(depth=1, dim=1)
!= region :: r2 = forward(depth=1, dim=2)
!= region :: robertsCross = r1*r2
!= stencil robertsCross :: a
\end{minted}
This is especially useful for common stencil patterns, such as Roberts cross,
as the region can be defined once and reused.
%%
\paragraph{Modifiers}
%%
Region specifications can be modified
by \emph{approximation} and \emph{multiplicity} information
(in \textit{spec} in Fig.~\ref{fig:syntax}).

The \texttt{readOnce} modifier enforces that no index appears more
than once (that is, its multiplicity is one). For example, in all of
the previous examples the \texttt{readOnce} modifier could be added,
\eg{}
%
\begin{minted}{fortran}
!= stencil readOnce, backward(depth=2, dim=1) :: a
b(i+1) = a(i) + a(i-1) + a(i-2)
\end{minted}
%
This specification would be invalid if any of the
array subscripts were repeated. This modifier provides a way to
rule out any accidental repetition of array subscripts.
The notion is similar to that of \emph{linearity} in linear type
systems~\cite{wadler1990linear}, where a value must be used
exactly once. We opt for the more informative and easily understood name
\texttt{readOnce}. This modifier is optional, so it need not
be present even if the stencil is linear.

In some cases, it is useful to give a lower and/or upper bound for a
stencil. This can be done using either the \term{atMost} or
\term{atLeast} modifiers. This is particularly useful in situations
where there is a non-contiguous stencil pattern, which cannot be expressed
precisely in our specification syntax. For example:
%
\begin{minted}{fortran}
!= stencil atLeast, pointed(dim=1)       :: a
!= stencil atMost, forward(depth=4, dim=1) :: a
b(i) = a(i) + a(i+4)
\end{minted}

\section{A model for specifications}
\label{sec:semantics}

We define a simple set-theoretic denotational model of the semantics
of our specification language. This has a number of purposes:
(1) it serves to explain the meaning of our specifications
(2) it is used in the inference and checking algorithms
(Section~\ref{sec:analysis}); (3) justifies an equational theory for
specifications in the next section; and (4) can be used to guide
correct implementations. The model domain is over sets of vectors
which we call \emph{index schemes}.

\begin{definition}
  An \emph{index scheme} is a vector of size $n$ (an $n$-vector) with
  values drawn from $\domainVal = \mathbb{Z} \cup \{\infty\}$.  Values
  in $\mathbb{Z}$ represent the offsets of neighbourhood indices
  (\ie{}, the constant integer $c$ in an indexing expression of the
  form $i \pm c$, Def.~\ref{def:neighbour},
  p.~\pageref{def:neighbour}) and $\infty$ is a wild-card
  representing both constant indices and un-specified indices. Throughout $u, v, w$ range over
  index schemes and $u_i$ denotes the $i^{th}$ element of $u$. We
  write $(\domainVal)^n$ for the set of all $n$-dimensional schemes.
\end{definition}

Indexing schemes give an abstract
representation of array subscripts in the source language.
Our model represents the meaning of stencil specifications
as indexing schemes, which explain the range of possible
indexing behaviours which are consistent with the specification.
The relationship between source language
terms and indexing terms
 is given precisely by the following function \textsf{schematic},
 which forms part of checking and inference (\S\ref{sec:analysis}).

\newcommand{\schemeOne}{\textsf{s}_I}
%\newcommand{\schemeOne}{\textsf{schema$^1_I$}}
\newcommand{\schemeN}{\textsf{schematic}_I}
%\newcommand{\schemeN}{\textsf{schema$^n_I$}}
\begin{definition}%[Schematic]
\label{def:schematic}
Let $\schemeOne$ be a partial function, parameterised by a set of
induction variables $I$, mapping expression syntax to
$\domainVal$ values:
%%
\begin{align*}
\schemeOne(e)
 & =  \begin{cases}
c & e \equiv i + c \;\; \wedge \;\; i \in I \\
-c & e \equiv i - c \;\; \wedge \;\; i \in I \\
\infty & \textit{$I \not\subseteq \textit{FV}(e)$} \\
\bot   & \textit{otherwise}
\end{cases}
\end{align*}
where $\textit{FV}(e)$ is the set of free variables in $e$. In the
first two cases, expressions that are a constant offset of an
induction variable (\ie{}, neighbourhood indices) are mapped to the
offset $c$ or $-c$ $\in \domainVal$. In the third case, if the
expression is not a neighbourhood index, but neither is it a relative
index (\ie{}, it is not defined in terms of an induction variable) then it
is mapped to $\infty \in \domainVal{}$. Otherwise, $\schemeOne(e)$
is undefined (returns $\bot$).

The partial function $\schemeN$, also parameterised by a set of
induction variables $I$, uses the above operation $\schemeOne$ to
 map from the syntax of indices to index schemes, \ie{} values of
$(\domainVal)^n$. It is defined on the syntax of indices $\bar{e} = (e_1, ..., e_n)$ as:
\begin{align*}
\schemeN(e_1, ..., e_n) & =
[\schemeOne(e_1) ... \schemeOne(e_n) ]
\quad \textit{iff} \;\; \forall x . \schemeOne(e_x) \neq \bot
\end{align*}
%
Thus, if each expression in an index is a neighbourhood or constant
index, then $\schemeN$ maps it to an indexing scheme.
\end{definition}

\begin{definition}[Generalisation]
Given two schemes $v, w$ of dimensionality $n$,
then $w$ is said to \emph{generalise} $v$
(or conversely $v$ \emph{specialises} $w$),
written $v \preceq_n w$, iff the following holds:
%%
\[
v \preceq_n w \;\; \Leftrightarrow  \;\;
  \forall i\!\in\!\{ 1, \ldots, n \} . \, (w_i = \infty \, \vee \, v_i = w_i)
\]
%%
That is, two schemes are related by the (reflexive and transitive) binary relation
$\preceq_n$ over $(\mathbb{Z}_{\infty})^n$ if in each dimension
the two schemes are equal or $w_i = \infty$
in place of some $\mathbb{Z}$ in $v_i$ for $v \preceq_n w$.
Thus, the element $\infty$ generalises any index in the same dimension.
%
\end{definition}

\begin{example}
Consider the following indexing schemes
corresponding to array indexing terms
$(i, j+1)$, $(1, j+1)$, $(3, j)$ and $(i*2, j)$:
%
\begin{align*}
\begin{array}{ll}
\schemeN{(i, j+1)} & =\vtwoh{0}{1} \\
\schemeN{(1, j+1)} & = \vtwoh{\infty}{1} \\
\schemeN{(3, j)} & = \vtwoh{\infty}{0} \\
\schemeN{(i*2, j)} & = \bot
\end{array}
\end{align*}
There is one generalisation between these
schemes, namely that $\vtwoh{0}{1} \preceq_2 \vtwoh{\infty}{1}$.
Note the last example is undefined as its first expression is a
non-neighbour relative index.
\end{example}

Our terminology of \emph{indexing schemes} and
\emph{generalisation} is by analogy
with the \emph{type schemes} and \emph{type generalisation} of
polymorphic types in ML~\cite{milner1978theory}. Our notion is
much more rigid however, as our schemes do not bind names.

\subsection{Denotational model}

The model is defined by the interpretation function $\interp{-}_n$
(where $n$ is the maximum dimensionality of the specification being
modelled) mapping closed\footnote{That is, we assume there are no
  occurrences of \textit{rvar} in a specification being modelled.  Any
  \emph{open} specification containing region variables can be made
  closed by straightforward syntactic substitution with a (closed)
  \textit{region}.} specifications to sets of
$n$-dimensional indexing schemes. The interpretation is overloaded on
\emph{regions} in Figure~\ref{fig:region-model} and on the top-level
of a specification \textit{spec} in Figure~\ref{fig:spatial-model}.

We first define some intermediate notations and definitions.

\begin{definition}A \emph{single-entry vector} of size $n$, denoted
$\singleEntry{r}{n}$, is a vector where the $r^{th}$ entry is $1$
and all others are $\infty$, \eg{}, $\singleEntry{1}{2} =
\vtwoh{1}{\infty}$ and $\singleEntry{2}{3} = \vthreeh{\infty}{1}{\infty}$.
Thus, a single-entry vector is an indexing scheme which generalises
all other schemes that have $1$ as their $r^{th}$ entry,
describing a neighbour offset of $1$.
\end{definition}

\begin{definition}
A \emph{zeroed single-entry vector} of size $n$, denoted
$\zeroEntry{r}{n}$, is a vector where the $r^{th}$ entry is $0$ and all others
are $\infty$, \eg{}, $\zeroEntry{1}{2} = \vtwoh{0}{\infty}$.
Similarly to the single-entry vector, zeroed single-entry vectors
generalise all other schemes
that have $0$ as their $r^{th}$, representing indices at
the ``origin'' in dimension $r$.
\end{definition}

\begin{figure}[t]
\vspace{-0.9em}
\input{region-model.tex}
%\textit{$a^m \in A$ means there are $n$ copies of $a$ in
%  the multi-set $A$}
\caption{Model of regions,
$\interp{-}_n : \textit{region} \rightarrow \mathcal{P}(\domainVal{}^n)$}
\label{fig:region-model}
\vspace{-1em}
\end{figure}

\noindent
The first four equations of Fig.~\ref{fig:region-model}
give the model of the region constants as sets of indexing
schemes. We illustrate with an example.

\begin{example}
For a $2$-dimensional stencil computation, then
\begin{align*}
& \interp{\stenCenS{2}{1}}_2 \\[-0.1em]
& \qquad = \; \{i\singleEntry{1}{2} \,|\, i \in \{\text{-}2,..., 2\}\!\setminus\!0 \}
  \cup  \{ \zeroEntry{1}{2} \,|\, \textit{true} \} \\[-0.2em]
& \qquad = \; \{-2\singleEntry{1}{2}, \, -\singleEntry{1}{2}, \,
  \singleEntry{1}{2}, \, 2\singleEntry{1}{2}\} \cup
  \{\zeroEntry{1}{2}\} \\[-0.2em]
& \qquad = \; \{\vtwoh{-2}{\infty}, \vtwoh{-1}{\infty}, \vtwoh{0}{\infty},
\vtwoh{1}{\infty}, \vtwoh{2}{\infty}\}
\end{align*}
Note the absorbing behaviour of $\infty$ with respect to
multiplication.
\end{example}
\noindent
The final two equations of Fig.~\ref{fig:region-model} give the models
of $\term{+}$ and $\term{*}$. For $\term{+}$ this is the
straightforward union of the models of subterms. The model of
$\term{*}$ takes the tensor $\otimes$ of subterm models. The tensor is
more involved; we give some informal intuition first.

% specifications correspond to a set of schemes

% r+s means an index must be generalised by a scheme in either r or s
%       and every scheme in r must have an index which specialises it
%       and every scheme in s must have an index which specialises it

% r*s means an index must be generalised by a scheme in r and a scheme in s
%       and every pairs of schemes in r and s must have an index which
%           specialises both
%

% r+s means indices must be consistent with either r or s
%           and there are indices why are generalised by r and s
% (there are indices which are consistent with every part of r or s)
% (every part of r and s independently have an index which is generalised by it)

% r*s means indices must be consistent simultaenously with both r and s
%           and there are indices which are generalised
%

\begin{figure}
\input{spatial-model.tex}
\caption{\textit{spec} model,
$\interp{-}_n : \textit{spec} \rightarrow \textsf{Mult}(\textsf{Approx}(\mathcal{P}(\domainVal{}^n)))$}
\label{fig:spatial-model}
\vspace{-0.75em}
\end{figure}

\paragraph{$\otimes$ of models}

The intuition behind $\otimes$ is that it takes all possible pairs of
indexing schemes, and treats each pair as lower and upper bounds of an
$n$-dimensional rectangle. From these lower and upper bounds,
the remaining vertices of the $n$-dimensional rectangle are
generated (like a bounding box) \eg{}:
%%
\begin{align*}
\{\vtwohs{1}{2}, \vtwohs{5}{6}\,\} \otimes \{\vtwohs{3}{4}\} =
\{\vtwohs{1}{2}, \vtwohs{3}{2}, \vtwohs{1}{4}, \vtwohs{3}{4},
   \vtwohs{3}{6}, \vtwohs{5}{4}, \vtwohs{5}{6}\,\}
\end{align*}
%%
The tensor takes a Cartesian product of the two models,
giving all pairs of schemes which are combined using the
\emph{pairwise permutation} written $\bowtie$ (defined later).
Pairwise permutation generates all possible interleavings of two
vectors. Some care is then taken over
how this interacts with $\infty$: pairwise permutation
corresponds to non-deterministic choice, in each dimension $i$,
between two values $u_i$ and $v_i$, but if either
is $\infty$ then the resulting pairwise permutation is
filtered out if the dimension $i$ has \emph{any other scheme
in the model where $u_i \neq \infty$}, \eg{}:
\begin{align*}
\{\vtwoh{1}{\infty}\} \otimes \{\vtwoh{3}{4}\} =
\{\vtwoh{1}{4}, \vtwoh{3}{4}\}
\end{align*}
%
where $\vtwoh{1}{\infty}$ and $\vtwoh{3}{\infty}$ have been filtered.

The notion of which dimensions $i$ have \emph{any scheme in the model
where $u_i \neq \infty$} is represented by the \emph{constrained
dimensions} intermediate function:
\begin{definition}%[Constrained dimensions]
For an $n$-dimensional model $M$, its \emph{constrained dimensions}
is the set of dimension identifiers for which at least one scheme $u \in M$ has
a non-$\infty$ value in that dimension. That is:
%
\begin{align*}
\effdims{M}{n} =
\bigcup_{u \in M} \{i \mid i \in \{1,\ldots,n\} \, \wedge \, u_i \neq
  \infty\}
\end{align*}
For example, $\effdims{\{\vtwoh{0}{\infty},\vtwoh{\infty}{\infty}\}}{2}
 = \{1\}$.
\end{definition}
\begin{definition}The tensor $\otimes_n$ of $n$-dimensional
models is defined:
\begin{alignat*}{2}
& M \otimes_n N =
  \Bigg\{x \; \Bigg| \;
    \parbox{5.2cm}{$(u, v) \in M \times N $ \\[0.1em]
                  $\;\; \wedge \;\, i \in \{1, \ldots, 2^n\} \, \wedge
                  \, x = (u \bowtie v)_i$ \\[0.1em]
                  $\;\; \wedge \;\, j \in (\effdims{M}{n} \cup
                   \effdims{N}{n})$\\[0.1em]
                   $\;\; \wedge \;\, x_j \neq \infty$
                  } \Bigg\}
\end{alignat*}
The first guard takes the Cartesian product of
the models and quantifies over all pairings $(u, v)$.
The second guard binds $j$ which ranges over the constrained
dimensions for $M$ unioned with the constrained dimensions for $N$.
That is, every $j$ corresponds to a dimension in which
either $M$ or $N$ has a scheme $u$ where $u_j \neq \infty$,
and thus is constrained.
The \emph{pairwise permutation} $u \bowtie v$
builds a $2^n \times n$ matrix
of all possible $n$-vectors generated by
non-deterministically picking for each $j^{th}$ entry either
$u_j$ or $v_j$ \eg{}
%
\begin{equation*}
\vthreeh{0}{1}{2} \bowtie \vthreeh{3}{4}{5} =
\left[
\setlength{\arraycolsep}{0.42em}
\begin{array}{cccccccc}
0 & 0 & 0 & 0 & 3 & 3 & 3 & 3 \\
1 & 1 & 4 & 4 & 1 & 1 & 4 & 4 \\
2 & 5 & 2 & 5 & 2 & 5 & 2 & 5
\end{array}
\right]
\end{equation*}
%
The $2^n$ unique choices for $\bowtie$ on $n$-vectors
corresponds to taking all bit-strings of length $n$ and
selecting from $u$ for $1$ and $v$ for $0$. The pairwise permutation
matrix is thus defined in terms of the logical
matrix $b$ where $b_{i,j}$ is the $j$-th bit of the integer
$i - 1$ as:
%
\begin{align*}
(u \bowtie v)_{i,j} & = b_{i,j} u_j + \neg b_{i,j} v_j
\end{align*}
%
The permuted schemes $x = (u \bowtie v)_i$ added to the model
$M \otimes N$ are only those which have do not have $\infty$ in
the constrained dimensions given by each model ($x_j \neq
\infty$). This filters out all permutations which have
$\infty$ in a constrained dimension position.
\end{definition}

\paragraph{Model of \textit{spec}}

Regions are modelled as sets of index
schemes, that is
$\interp{\textit{region}}_n \in \mathcal{P}(\domainVal{}^n)$.
Figure~\ref{fig:spatial-model} defines the top-level model of the
\textit{spec} syntax, which includes the multiplicity and
approximation modifiers. The domain of this interpretation
is $\textsf{Mult}(\textsf{Approx}(\mathcal{P}(\domainVal{}^n)))$
where $\textsf{Mult}$ and $\textsf{Approx}$ are labelled variants
labelling our set-based model with additional information.

\begin{definition} \textsf{Mult}
and \textsf{Approx} are parametric labelled variant types
with injection function given by their definition:
%
\begin{align*}
\textsf{Mult} \, a & = \textsf{mult} \, a \,\mid\, \textsf{only} \, a \\
\textsf{Approx} \, a & = \textsf{exact} \, a \,\mid\, \textsf{lower} \, a \,\mid\,
\textsf{upper} \, a \,\mid\, \textsf{both} \, a \, a
\end{align*}
\eg{}, $\mathsf{lower}$ is an injection
$\mathsf{lower} : a \rightarrow \mathsf{Approx} \, a$ etc. The
$\textsf{Mult}$ type corresponds to the presence or absence of the
\term{readOnce} modifier as shown in Figure~\ref{fig:spatial-model}.
The $\textsf{Approx}$ type corresponds to the presence or absence of
the spatial approximation modifier, with \textsf{exact} when there is
no such modifier and \textsf{lower} and \textsf{upper} for
\term{atLeast} and \term{atMost}. Later, it will be useful to model
pairs of lower and upper bounds provided by \textsf{both} but which is
not used
here.% These are used for labelled models, and other components later in.
\label{def:mult-and-approx}
\end{definition}

\section{Equational theory}
\label{sec:eqs}

Specifications are subject to a set of equations which can be used to
understand when specifications are equivalent. Furthermore,
specifications are subject to a notion of approximation. We outline
the equational theory and notion of approximation here.

\begin{figure}
\vspace{-0.3em}
\begin{align*}
\setlength{\arraycolsep}{0.05em}
\begin{array}{c}
%\framebox{$\textit{region} \equiv \textit{region}'$} \\[0.6em]
\begin{array}{rll}
(\textsc{F\, \texttt{+} \,F}) \;\; &
\stenFwdSR{n \, \textsf{max} \, m}{d}{p \vee q} \\
 \equiv \; & \stenFwdSR{n}{d}{p} \; \texttt{+} \; \stenFwdSR{m}{d}{q} \\[0.6em]
%
%
(\textsc{B\, \texttt{+} \,B}) \;\; &
\stenBwdSR{n \, \textsf{max} \, m}{d}{p \vee q} \\
 \equiv \; & \stenBwdSR{n}{d}{p} \; \texttt{+} \; \stenBwdSR{m}{d}{q} \\[0.6em]
%
%
(\textsc{C\, \texttt{+} \,C}) \;\; &
\stenCenSR{n \, \textsf{max} \, m}{d}{p \vee q} \\
\equiv \; & \stenCenSR{n}{d}{p} \; \texttt{+} \; \stenCenSR{m}{d}{q} \\[0.6em]
%
%
(\textsc{C\, \texttt{+} \,F}) \;\; & \stenCenSR{n}{d}{p \vee q} \\
\textit{$m \leq n$} \; \equiv \; & \stenCenSR{n}{d}{p} \; \texttt{+} \;
                      \stenFwdSR{m}{d}{q} \\[0.6em]
%
%
(\textsc{C\, \texttt{+} \,B}) \;\; &
\stenCenSR{n}{d}{p \vee q} \\
\textit{$m \leq n$} \; \equiv \; & \stenCenSR{n}{d}{p} \; \texttt{+} \;
                      \stenBwdSR{m}{d}{q} \\[0.6em]
%
%
(\textsc{B\, \texttt{+} \,F}) \;\; &
\stenCenSR{n}{d}{p \vee q} \\
\equiv \; & \stenFwdSR{n}{d}{p} \; \texttt{+} \; \stenBwdSR{n}{d}{q}
  \\[0.6em]
%
%
(\textsc{P\, \texttt{+} \,F}) \;\; &
\stenFwdS{n}{d} \\
\equiv \; & \stenReflSR{d} \; \texttt{+} \; \stenFwdSR{n}{d}{p} \\[0.6em]
%
%
(\textsc{P\, \texttt{+} \,B}) \;\; &
\stenBwdS{n}{d} \\
\equiv \; & \stenReflSR{d} \; \texttt{+} \; \stenBwdSR{n}{d}{p} \\[0.6em]
%
%
(\textsc{P\, \texttt{+} \,C}) \;\; &
\stenCenS{n}{d} \\
\equiv \; & \stenReflSR{d} \; \texttt{+} \; \stenCenSR{n}{d}{p} \\[0.6em]
%
%
(\textsc{npoint}) \;\; & (R \, \term{*} \, S^\irreflS) \, \term{+} \,
                         (R^\irreflS \, \term{*} \, S) \;
\equiv \; R \, \term{*} \, S \\[0.4em]
%
(\textsc{$\ast$Dim}) \;\; & R^d \, \term{$\ast$} \, S^d \equiv R^d
                             \, \term{+} \, S^d \\[0.4em]
%
(\texttt{+}\textsc{idem}) \;\; & S \; \texttt{+} \; S \; \equiv \; S \\[0.4em]
%
(\texttt{+}\textsc{comm}) \;\; & S \; \texttt{+} \; T \; \equiv \; T \;
                       \texttt{+} \; S \\[0.4em]
%
(\texttt{+}\textsc{assoc}) \;\; & R \; \texttt{+} \; (S \; \texttt{+} \; T) \; \equiv \; (R \;
                       \texttt{+} \; S) \; \texttt{+} \; T \\[0.4em]
(\texttt{*}\textsc{comm}) \;\; & S \; \texttt{*} \; T \; \equiv \; T \;
                       \texttt{*} \; S \\[0.4em]
%
(\texttt{*}\textsc{assoc}) \;\; & R \; \texttt{*} \; (S \; \texttt{*} \; T) \; \equiv \; (R \;
                       \texttt{*} \; S) \; \texttt{*} \; T \\[0.4em]
(\textsc{dist}) \;\; & R \; \texttt{*} \; (S \; \texttt{+} \; T) \; \equiv \; (R \;
                       \texttt{*} \; S) \; \texttt{+} \; (R
                       \; \texttt{*} \; T)
\end{array}
\end{array}
\end{align*}
\caption{Equations on specifications (\emph{regions})}
\vspace{-0.5em}
\label{fig:equations}
\end{figure}

Figure~\ref{fig:equations} lists the equational theory for
specifications given by the binary relation $\equiv$
on \emph{region} terms. The equations use abbreviations
\term{point}, \term{fwd}, \term{bwd}, \term{cen}
for \term{nonpointed}, \term{forward}, \term{backward},
\term{centered} respectively. Furthermore, the abbreviations include syntactic sugar
for quantifying over the absence or presence of the \irrefl{}
attribute on region constants as a boolean. For example,
\term{fwd} syntactic sugar is defined:
%%
\begin{align*}
\hspace{-0.25em}
& \stenFwdSR{n}{d}{r}
\; := \; \\[-0.4em]
& \qquad \begin{cases}
\stenFwd{n}{d}{, \irrefl} & \;\;\textit{if} \;\, \neg \, r \\[-0.1em]
\stenFwd{n}{d}{}  & \;\;\textit{if} \;\,  r
\end{cases} \\
\textit{or} \;\;
& \stenFwdS{n}{d} := \stenFwd{n}{d}{}
\end{align*}
%
and similarly for \term{bwd} and \term{cen}.

Stencil specifications in \textit{spec} are considered
equal when they have the same modifiers and $\equiv$ equates
their regions, \ie{}:
\begin{align*}
(\textit{region} \equiv \textit{region'})
 \Rightarrow \quad\;\; & \texttt{stencil} \; [\textit{approx},] \; [\textit{mult},] \;
\textit{region} \\[-0.4em]
\quad \equiv \; & \texttt{stencil} \; [\textit{approx},] \;
            [\textit{mult},] \; \textit{region}'
\end{align*}
%

\paragraph{Overlapping regions with +}
The first nine
equations (with labels of the form $\ldots \texttt{+} \ldots$)
explain the notion of \emph{region overlapping} with sum
\term{+}. In the first six, $p$ and $q$ range
over booleans denoting the absence (true) or presence (false)
of \irrefl{}. These are combined via disjunction. For example,
the rule $(\textsc{B} \, \term{+} \,
\textsc{F})$ explains that the sum of \term{backward}
and \term{forward} regions in the same dimension and the same
depth is equivalent to a \term{centered} region of that depth.
If the \term{forward} part is non-pointed ($p =
\textit{false}$) but \term{backward} is pointed ($q =
\textit{true}$) then their sum is equal to a \term{centered}
region with no \irreflS{} modifier by the disjunction ($p \vee q =
\textit{true}$), \ie{}:
%%
\begin{align*}
& \stenCenS{n}{d}  \\[-0.2em]
\equiv \;\; & \stenFwdSR{n}{d}{\irrefl} \; \texttt{+} \; \stenBwdS{n}{d}
\end{align*}
%%
That is, the \term{fwd} region is non-pointed describing code which is missing
a zero-offset neighbourhood index (\eg{}, missing $\texttt{a(i)}$), but the addition of
the \term{bwd} has no non-pointed attribute and thus provides the
missing zero offset index.

The equation (\textsc{point}) provides an annihilation of
\irrefl{} using a further syntactic sugar.
We write $R^{\irreflS}$ to denote a region constant with an
non-pointed attribute, \eg{},
%%
\begin{align*}
\stenFwdS{n}{d}^{\irreflS}  := \,
\stenFwdSR{n}{d}{\irrefl}
\end{align*}
%%
The left-hand side of (\textsc{point}) has a sum of
products of two regions $R$ and $S$,
where in each component of the sum one of $R$
or $S$ has the non-pointed attribute. These cancel to give
just $R \, \term{*} \, S$.

\paragraph{Interaction between \term{+} and \term{*}}

The last seven equations (from (\textsc{\term{*}Dim}) inclusive), explain
the algebraic behaviour of \term{+} and \term{*}.  Together, we see
that \term{+} is an associative, commutative, idempotent binary
operator that distributes with \term{*}, which is associative and
commutative. This distribution of \term{*} over \term{+} is key as it
is used internally to give a normal form for stencil specifications
akin to Disjunctive Normal Form (DNF) (where \term{+} is
disjunction and \term{*} is conjunction). We revisit this
normalisation in \S~\ref{sec:inf-step4}.

The equation (\textsc{$\ast$Dim}) uses syntactic sugar
where $R^d$ and $S^d$ denote regions whose dimension is $d$.
This equation shows that a product of two regions on
the same dimension is equal to their addition.

The equational theory provides not only an axiomatic semantics
for the language, which the user can use to rewrite their
specification, but it is also used extensively in the inference
procedure (\S~\ref{subsec:inference}) to simplify specifications.

\begin{theorem}[Soundness of equational theory]
\begin{equation*}
\forall S, T, n . \quad
S \equiv T \; \Rightarrow \;
\interp{S}_n = \interp{T}_n
\end{equation*}
\end{theorem}

\subsection{Approximation}

Figure~\ref{fig:inequations} defines a notion of sub-specifications
via the relation $<:^r$ on \textit{region}
syntax and then $<:$ on \textit{spec} syntax inductively.
Both relations are reflexive and transitive (we omit
these equations) and congruent
with respect to \term{+}, \term{*}, \term{atLeast}, \term{atMost},
and \term{readOnce}. The \trule{\textsc{eq}} rule connects region
equations with inequations. The \trule{\textsc{over}} rule explains
the notion of spatial over-approximation via overlapping regions.
The \trule{\textsc{rep}} allows the modifier that
enforces that every index is read just once to be dropped
as an approximation; \trule{\textsc{bound}} gives
a connection between the \term{atLeast} and \term{atMost}
approximation modifiers.

\begin{theorem}[Soundness of approximation]
\begin{equation*}
\forall S, T, M, n . \quad
S <: T \; \wedge (\consA{M}{\interp{S}_n})
\Rightarrow (\consA{M}{\interp{T}_n})
\end{equation*}
\end{theorem}

\begin{figure}[t]
\begin{align*}
\hspace{-0.5em}
\begin{array}{c}
%%
%% INEQUATIONS ON REGION
%%
%\framebox{$\textit{region} <: \textit{region}'$} \\[0.85em]
%
%\dfrac{}{R <:^r R}\trule{\textsc{refl}} \qquad \dfrac{P <:^r R \quad R <:^r Q}{P <:^r
%  Q}\trule{\textsc{trans}}
%\qquad
%\\[1.5em]
\vspace{-0.4em}
\setlength{\arraycolsep}{0.1em}
\dfrac{\hspace{3em} m \leq n \qquad\qquad r \implies s \hspace{3em}}
{\begin{array}{rl}
\stenFwdSR{m}{d}{r} & <:^r \stenFwdSR{n}{d}{s}
%& \;\, ds \subseteq es \wedge n \leq m
\\
\wedge \; \stenBwdSR{m}{d}{r} & <:^r \stenBwdSR{n}{d}{s}
%& \;\, ds \subseteq es \wedge n \leq m
\\
\wedge \; \stenCenSR{m}{d}{r} & <:^r \stenCenSR{n}{d}{s}
%& \;\, ds \subseteq es \wedge n \leq m
\\
\wedge \; \stenBwdSR{m}{d}{r} & <:^r \stenCenSR{n}{d}{s} \\
\wedge \; \stenFwdSR{m}{d}{r} & <:^r \stenCenSR{n}{d}{s} \\[1.2em]
\end{array}}\trule{\textsc{over}} \\[-0.2em]
\hspace{-0.5em}
\dfrac{R_1 <:^r S_1 \quad R_2 <:^r S_2}
      {R_1 \, \texttt{+} \, R_2 <:^r S_1 \, \texttt{+} \, S_2}
\trule{\textsc{cong}\texttt{+}}
\dfrac{R_1 <:^r S_1 \quad R_2 <:^r S_2}
      {R_1 \, \texttt{*} \, R_2 <:^r S_1 \, \texttt{*} \, S_2}
\trule{\textsc{cong}\texttt{*}} \\[1.5em]
%%
%% INEQUATIONS ON SPEC
%%
%\framebox{$\textit{spec} <: \textit{spec}'$} \\[0.8em]
%
%
\dfrac{R <:^r S}{\term{atLeast}  S <: \term{atLeast}
  R}\trule{\texttt{L}}
\;\;
\dfrac{R <:^r S}{\term{atMost}  R <: \term{atMost}
  S}\trule{\texttt{M}} \\[1.5em]
\dfrac{R <:^r S}{R <: \term{atMost} \, S}\trule{\textsc{grow}} \quad
\dfrac{R <:^r S}{S <: \term{atLeast} \, R}\trule{\textsc{shrink}} \quad
 \\[1.5em]
%%
\dfrac{\mathsf{dims}(R) \cap \mathsf{dims}(S) = \emptyset}{R \ast S <: R}
\trule{\textsc{Gen$\ast$}}
\quad
\dfrac{R \equiv S}{R <: S}\trule{\textsc{eq}}
%%
\\[1.5em]
\dfrac{R <: S}{\term{readOnce} \, R <: \term{readOnce} \,
  S}\trule{\textsc{cong}\texttt{R}}
\;\;
\dfrac{}{\term{readOnce} \, R <: R}\trule{\textsc{rep}}
%\dfrac{}{R <: \term{atMost} R}
%\quad
%\dfrac{}{R <: \term{atLeast} R}\\[1.5em]
\end{array}
\end{align*}
\caption{Inequations $<:^r$ on \textit{regions} and
$<:$ on \textit{specs} (omitting reflexivity
and transitivity). }
\label{fig:inequations}
\vspace{-0.8em}
\end{figure}

\section{Analysis, checking, and inference}
\label{sec:analysis}

\noindent
We outline here the procecures for checking the conformance
of source code against specifications (\S~\ref{subsec:checking})
and for infering specifications from code (\S~\ref{subsec:inference}).
Both rely on the analysis of source code, converting array indexing
expressions into sets of index schemes. We outline this analysis
first (\S~\ref{subsec:analysis}). Note that the analysis procedure
can be made arbitrarily more complex and wide-ranging independent
of the checking and inference procedures.

\subsection{Analysis of array-based code}
\label{subsec:analysis}

\newcommand{\neigh}{\textsf{neigh}}

\paragraph{Syntax helper functions}
The function \rhsExp{} maps source statements to a
set of expressions which occur in right-hand positions (\ie{}, not the
target of an assignment). The function \var{} maps expressions to a
set of the variables used in right-hand positions.

\paragraph{Analysis pre-requisites}
The analysis builds on standard program analyses:
%
\textbf{(1)} basic blocks (CFG);
\textbf{(2)} induction variables per basic block;
\textbf{(3)} (interprocedural) data-flow analysis, providing a \emph{flows to}
  graph (as shorthand, the function
  $\mathsf{flowTo}$ is used, implicitly parameterised by this graph,
  mapping an expression to the set of all expressions
  with forwards data-flow to this expression, based on the transitive
  closure of the flows graph);
\textbf{(4)} type information per variable, where we use the predicate
\arrayTy{} to classify variables of array type.

\paragraph{Core analysis}
For each assignment statement \textit{assg} within a loop whose left-hand side is an array
subscript on neighbourhood indices, a finite map is computed mapping
array variables to \emph{multisets} of indexing schemes.
This is captured by the function $\textit{analyse}$:
%
\begin{align*}
\hspace{-0.5em}
\begin{array}{lr}
\textit{analyse}(a(\overline{e_1}) = e_2)
 := \qquad\qquad\qquad & \textit{where} \; \neigh(\overline{e_1}) \\[0.3em]
\multicolumn{2}{l}{\qquad \bigcup\{a' \mapsto \{
\textsf{schematic}(\bar{e})\} \mid a'(\bar{e}) \leftarrow \mathsf{flowsTo}(e_2),
  \arrayTy(a')\}}
\end{array}
\end{align*}
%
The predicate $\neigh$ classifies neighbourhood indices, based on
the induction variables of the loop (implicit here).
Thus, we focus on assignments to an array subscript where the LHS
index $\overline{e_1}$ is a neighbourhood index.  For
all array subscripts that flow to the right-hand side of this
statement, a finite map is constructed, mapping each array variable
to a \emph{multiset} of indexing schemes for its subscripts, computed
with \textsf{schematic} (Defn.~\ref{def:schematic},
p.~\pageref{def:schematic}).

Note that \textsf{schematic} is undefined ($\bot$) if
$\bar{e}$ contains relative indices which are not neighbourhood
indices. The \textit{analysis} function feeds into the
\textit{modelise} function which filters out those arrays
which contain undefined indexing schemes and adds multiplicity
information:
%
\begin{align*}
& \mathit{modelise} (\textit{assg})
= a \mapsto
\begin{cases}
\bot & \bot \in M \\
\textsf{only}(|M|) & \forall u \in M . \#(M, u) = 1 \\
\textsf{mult}(|M|) & \forall u \in M . \#(M, u) = n
\end{cases} \\
& \quad \textit{where} \; M = \mathit{analyse}(\textit{assg}) (a)
\end{align*}
where $\#(M, u)$ denotes the multiplicity of an element $u$ of the
multiset $M$ and $|M|$ denotes the set of elements in the multiset
$M$.

Thus, the analysis maps source code to index schemes, which is then
augmented with multiplicity information by $\mathit{modelise}$,
\ie{} values in $\mathsf{Mult}(\mathcal{P}((\domainVal{})^n))$.

If the LHS $\overline{e_1}$ contains non-$0$ neighbour offsets
then all RHS schemes are \emph{relativised} by
$\textsf{schematic}(\overline{e_1})$. For example,
if \texttt{a(i+1) = b(i)} then relativisation of the RHS by the LHS
produces an analysis which is equivalent to the analysis of \texttt{a(i) = b(i-1)}.
\dnote{Somewhere we need a discussion about syntactic vs semantic
analysis...}

\begin{example}
\label{exm:checking}
We demonstrate the analysis and modelise procedures over the
following example program which computes the mean value
of a five-point stencil:
%%
\begin{minted}{fortran}
do i = 1, (n-1)
  do j = 1, (m-1)
    x = a(i-1, j)+a(i+1, j); y = a(i, j-1)+a(i, j+1)
    b(i, j) = (a(i, j) + x + y) / 5.0
  end do
end do
\end{minted}
%
The \textit{analyse} function matches on
line 4, with the following set for $\textsf{flowsTo}(\texttt{a(i, j) + x +
  y})$:
%
\begin{align*}
\{\texttt{a(i-1, j)}, \texttt{a(i+1, j)}, \texttt{a(i, j-1)},
  \texttt{a(i, j+1)}, \texttt{a(i, j)}\}
\end{align*}
Subsequently the result of \textit{analyse} on line 4 yields the map:
\begin{align*}
\texttt{a} \mapsto \{\vtwoh{-1}{0}^1, \vtwoh{1}{0}^1,
          \vtwoh{0}{-1}^1, \vtwoh{0}{1}^1, \vtwoh{0}{0}^1\}
\end{align*}
%
where multiplicities are denoted by superscripts. The \emph{modelise}
function then yields the following since all multiplicities in the model are one:
%%
\begin{align*}
\texttt{a} \mapsto \mathsf{only}(\{\vtwoh{-1}{0}, \vtwoh{1}{0},
          \vtwoh{0}{-1}, \vtwoh{0}{1}, \vtwoh{0}{0}\})
\end{align*}
%%
\end{example}

%Given an assignment $\textit{assg}$ to an array subscript in the source language
%and an array variable $a$, then $\mathit{modelise}(\textit{assg}, a)$ performs
%an analysis to find all array subscripts of $a$ converting them to
%index schemes. For the moment, we leave this analysis abstract
%as the function $\emph{analyse}$ in the following definition of $\emph{modelise}$:
%

\subsection{Specification checking}
\label{subsec:checking}

In checking mode, the access pattern of stencil computations in the
source language is verified against any specification associated with
them in the code. The model of
\S~\ref{sec:semantics} guides this process.  Checking
proceeds by generating a model from a specification and generating
a model from the source code, comparing them for
\emph{consistency}. The notion of consistency is built out of index scheme
\emph{generalisation} (Sect.~\ref{subsec:schemes-and-gen}).

\begin{definition}[Checking]
Let $a$ be an array variable and let
$S \in \textit{spec}$ be a specification associated to $a$
via \texttt{!= stencil $S$ :: a}. A stencil kernel
 $v(\overline{e_1}) = e_2$ in the source language
conforms to $S$ if the predicate $\textbf{check}_n(S, v(\overline{e_1}) = e_2, a)$ holds,
defined:
\begin{align*}
& \textbf{check}_n(S, \textit{assg}, a) =
  \consSub{n}{\textit{modelise}(\textit{assg})(a)}{\interp{S}_n}
&
\end{align*}
Thus, checking proceeds by using \textit{modelise}
(\S~\ref{subsec:analyis}) to extract index schemes from source code
augmented with multiplicitly for the whole set of schemes.
This is compared with the model of the specification $\interp{S}_n$ by
the consistency predicate $\consAName$.
\end{definition}

\begin{definition}
Consistency relates models of source code to models
of specifications (which are bounded) via two predicates:
\begin{align*}
\consName_n & \subseteq \mathcal{P}(\domainVal{}^n)
\times \textsf{Approx}(\mathcal{P}(\domainVal{}^n)) \\
\consAName_n & \subseteq \textsf{Mult}(\mathcal{P}(\domainVal{}^n))
\times \textsf{Mult}(\textsf{Approx}(\mathcal{P}(\domainVal{}^n)))
\end{align*}
%
where $\consName$ relates models without multiplicity.
Figure~\ref{fig:consistency} defines the two predicates, where for
brevity we make implicit the subscript $n$ defining the
dimensionality.  We denote sets of indexing schemes
$\mathcal{P}((\domainVal{})^n)$ that come from a specification model
by $M$ and from source code model by $N$.

Rules (\emph{up}) and (\emph{low}) are the core rules relating models of
specifications and models of code. The (up) rule
considers the case where the specification is
an \term{atMost} upper-bound approximation.
An upper bound specification means that for every scheme in the code
model, modelling a source-level index, $\forall u \in C$ there exists
a generalising scheme in the spec model $\exists v \in S$, \ie{}, $u
\preceq v$. That is, every source-level index scheme is generalised by the specification.
The (down) rule provides the case of
\term{atLeast} lower-bound specifications. Consistency is dual to the above:
for every scheme in the specification model $\forall v \in S$ there
exists a scheme in the code model $\exists u \in C$ such that $u
\preceq v$. That is, every specification scheme is generalised by some source
index scheme.

The (\emph{up}) and (\emph{down}) rules are combined for an \textsf{exact}
specification (\textit{exact}) such that a specification's model
provides both a consistent lower and consistent upper bound on
for the source code. Thus, for every scheme in the code model there
is a scheme in the spec model that generalises it, and for every
scheme in the spec model there is a scheme in the source model that
specialises it.  The (\textit{both}) rule is similar, decomposing
a specification which has both a lower and upper bound into
consistency of these bounds.

Rules $(1)$ and $(*)$ extend the \consName{}
predicate to $\consAName{} \subseteq
\textsf{Mult}(\mathcal{P}(\domainVal{}^n))
\times \textsf{Mult}(\textsf{Approx}(\mathcal{P}(\domainVal{}^n)))$
which inludes multiplicity information. Both rules
state that consistent models remain consistent if they
have the same multiplicity. Relatedly, rule $(1 < *)$ ensures
that absence of a \texttt{readOnce} modifier in the specification
does not prohibit a set of unique stencil accesses in the
code from being consistent.
\end{definition}

\begin{figure}[t]
\begin{align*}
\hspace{-0.5em}
\begin{array}{c}
%%
\trule{up}\dfrac{\forall u \in M, \exists v \in N .  \, u \preceq v}
      {\cons{M}{\textsf{upper}(N)}}
\,
\trule{low}\dfrac{\forall v \in N, \exists u \in M . \, u \preceq v}
      {\cons{M}{\textsf{lower}(N)}} \\[1.5em]
%%
\trule{exact}\dfrac{\cons{M}{\textsf{lower}(N)} \qquad \cons{M}{\textsf{upper}(N)}}
  {\cons{M}{\textsf{exact}(N)}} \\[1.5em]
%%
\trule{both}\dfrac{\cons{M}{\textsf{lower}(M_1)} \qquad \cons{M}{\textsf{upper}(M_2)}}
  {\cons{M}{\textsf{both}(M_1, M_2)}}\\[1.5em]
%%
\trule{$1$}
\dfrac{\cons{M}{N}}{\consA{\textsf{once}(M)}{\textsf{once}(N)}}
\;\;
\trule{$*$}
\dfrac{\cons{M}{N}}{\consA{\textsf{mult}(M)}{\textsf{mult}(N)}}
  \\[1.5em]
\trule{$1 < *$}
\dfrac{\cons{M}{N}}{\consA{\textsf{once}(M)}{\textsf{mult}(N)}}
\end{array}
\end{align*}
\vspace{-0.5em}
\caption{Consistency of models (code vs. specification)}
\label{fig:consistency}
\vspace{-0.9em}
\end{figure}

\begin{example}
  We give an extended example of verifying a stencil computation
  against its specification. We consider a more general, but still
  consistent, version of the \textit{five-point} stencil specification
  inferred previously in \S~\ref{subsec:inference}:
%
\begin{minted}[breakindent=8.1em]{fortran}
!= region :: fivepoint = centered(depth=1,dim=1) + centered(depth=1,dim=2)
!= stencil readOnce, fivepoint :: b
a(i,j) = b(i,j-1)+b(i,j)+b(i,j+1)+b(i-1,j)+b(i+1,j)
\end{minted}
%
We refer to line three as $l3$ for brevity,
which is checked via 
$check(({\textcolor{darkgreen}{\texttt{readOnce, fivepoint}}}),l3,\texttt{b})$
with steps:
\begin{enumerate}
%
\item Generate the model from the specification as in
  \S~\ref{sec:semantics}.
%
\begin{align*}
 N_s= \{  \vtwo{-1}{\infty},
          \vtwo{0}{\infty},
          \vtwo{1}{\infty},
          \vtwo{\infty}{-1},
          \vtwo{\infty}{0},
        \vtwo{\infty}{1} \}  \\
\interp{{\textcolor{darkgreen}{\texttt{readOnce, fivepoint}}}}_2 =
  \textsf{once}(\textsf{exact} (N_s))
\end{align*}
%
\item Obtain the code model via $\textit{modelise}$, giving:
%
\begin{align*}
M_c = & \{
    \vtwo{0}{-1},
    \vtwo{0}{0},
    \vtwo{0}{1},
    \vtwo{-1}{0},
    \vtwo{1}{0}
  \}
\\
\textit{modelise}(l3, \texttt{b}) = & \textsf{once} (M_c)
\end{align*}
%
\item By the definition of consistency (Fig.~\ref{fig:consistency})
 $\textsf{once}(\textsf{exact}(N_s))$ is consistent with
  $\textsf{once}(M_c)$ when:
  \begin{align*}
    & \consSub{2}{\textsf{up}(N_s)}{M_c} \; 
    \wedge \; \consSub{2}{\textsf{low}(N_s)}{M_c}
  \end{align*}
%
\item This holds by the following sets of scheme
  generalisations: for the upper bound, $\forall u \in M_c,
  \exists v \in N_s . u \preceq_2 v$:
  \begin{align}
    \hspace{-1em}
    \begin{array}{l}
   \hspace{-1.5em}
     \vtwoh{0}{-1} \preceq_2 \vtwoh{0}{\infty} \quad
     \vtwoh{0}{0} \preceq_2 \vtwoh{0}{\infty} \quad
     \vtwoh{0}{1} \preceq_2 \vtwoh{0}{\infty} \\[0.5em]
        \hspace{-1.5em}\vtwoh{-1}{0} \preceq_2 \vtwoh{\infty}{0} \quad
     \vtwoh{1}{0} \preceq_2 \vtwoh{\infty}{0}
     \end{array}
  \label{eq:check-example-upper}
  \end{align}
  %
  and for the lower bound: $\forall v \in N_s, \exists u \in M_c . u
  \preceq_n v$:
  \begin{align}
    \hspace{-1em}
    \begin{array}{l}
     \hspace{-1.3em}\vtwoh{-1}{0} \preceq_2 \vtwoh{-1}{\infty} \quad
     \vtwoh{0}{0} \preceq_2 \vtwoh{0}{\infty} \quad
     \vtwoh{1}{0} \preceq_2 \vtwoh{1}{\infty}  \\
     \hspace{-1.3em}\vtwoh{0}{-1}  \preceq_2 \vtwoh{\infty}{-1} \quad
     \vtwoh{0}{0}  \preceq_2 \vtwoh{\infty}{0} \quad
     \vtwoh{0}{1}  \preceq_2 \vtwoh{\infty}{1}
     \end{array}
    \label{eq:check-example-lower}
  \end{align}
  Since the consistency predicate is satisfied, the array
  accesses flowing to line three in the stencil computation are valid
  with respect to the {\textcolor{darkgreen}{\texttt{fivepoint}}} specification.
\end{enumerate}
\vspace{-0.7em}
\end{example}
\begin{example}
Consider an erroneous user modification to $l3$ with an
additional term \texttt{b(i+2,j+2)} such that $l3'$ is:
%
\begin{minted}[firstnumber=3]{fortran}
a(i,j) = b(i,j-1) + b(i,j) + b(i,j+1) + b(i-1,j) + b(i+1,j) + b(i+2,j+2)
\end{minted}
%
The checking procedure proceeds in the same way as the first three
steps of the previous example, except that the result of
$\mathit{modelise}(l3',\texttt{b})$ generates an additional scheme
$\vtwoh{2}{2}$. Since the spec model has not changed and the code
model is has only the additional scheme $\vtwoh{2}{2}$, the generalisations for
the lower bound \eqref{eq:check-example-lower}
of step 4 above still holds since its consistency condition
universally quantifies over the \emph{spec model}. However, for
the upper bound, the consistency condition universally quantifies
over scheme $u$ of the \emph{code model} for which there exists
a scheme $v \in N_s$ such that $u \preceq_2 v$. Thus, we require a $v
\in N_s$ such that $\vtwoh{2}{2} \preceq_2 v$ but no such generalising
scheme existing in the spec model. Hence the model fails to hold as an upper bound.

Note that if the specification were as follows:
%
\begin{minted}{fortran}
!= stencil atLeast, readOnce, fivepoint :: b
\end{minted}
%
then consistency holds as it no longer requires the failing upper-bound case:
$\consSub{2}{\textsf{mult}(\textsf{up}(N_s))}{\textsf{mult}(M_c)}$.
 This makes sense semantically since the region ${\textcolor{darkgreen}{\texttt{fivepoint}}}$
is indeed contained in the regions touched by $\texttt{b}$ in $l3'$.
\end{example}

\dnote{Build this into the definitions earlier I think}
One way the implementation may deviate from the model is elimination of all $\infty$
offset vector that occurs when all indices are in non-neighbourhood form \eg{}
\mintinline{fortran}{b(42,0)}. The only generalisation scheme that can unify
with this pattern is the most general one, meaning it requires outright
rejection of an otherwise fine stencil computation. See the following example,
where not discarding all $\infty$ access pattern prevents verification:
%
\begin{minted}{fortran}
!= stencil forward(dim=1,depth=1) :: a
a(i) = a(i) + a(i+1) + a(1)
\end{minted}

\subsection{Specification inference}
\label{subsec:inference}

We illustrate each step of inference using the
 five-point stencil program from Example~\ref{exm:checking} (p.~\pageref{exm:checking}).

\subsubsection{Step 1: Data access analysis}
\label{sec:inf-step1}

The first step performs the
analysis of \S~\ref{subsec:analysis} to produce a
model of the code in terms of indexing schemes.
The \emph{analyse} and \emph{modelise} procedures before yielded the finite map:
%
\begin{align*}
\texttt{a} \mapsto \mathsf{only}(\{\vtwoh{-1}{0}, \vtwoh{1}{0},
          \vtwoh{0}{-1}, \vtwoh{0}{1}, \vtwoh{0}{0}\})
\end{align*}
%

\subsubsection{Step 2: Coalesce contiguous schemes into spans}
\label{sec:inf-step3}

Let $U$ range over the finite maps computed by \textit{modelise}.  For
each array variable $a \in \mathsf{dom}(U)$, the algorithm constructs next a set of
$n$-dimensional rectangles covering all contiguous groups of schemes
in $U(a)$.

\newcommand{\spanOp}{\textsf{spans}}

\begin{definition}[Spans]
  A \emph{span} represents an $n$-dimensional rectangle
  as a pair of $n$-dimensional vectors (or
  represented as a $2 \times n$ matrix) with values drawn from
  $\domainVal{}$. This gives the co-ordinates of a lower-bound vertex
  (first component) and the upper-bound (second component). We write
  spans as $\vect{x} = [\vect{x}^L \; \vect{x}^U]$, where $\vect{x}^L$
  and $\vect{x}^U$ give the lower and upper bounds.  In
  the following, $\vect{x}, \vect{y}$ range over spans.  For a span
  $\vect{x}$, we write $\vect{x}_{i:j}$ for the sub-span comprising the
  subvectors $[\vect{x}^L_{i:j} \; \vect{x}^U_{i:j}]$ from entry $i$
  to entry $j$ (inclusive).
\end{definition}

\noindent
Each indexing scheme $u \in U(a)$ is converted to a \emph{unit} span
$u \mapsto [u \; u]$, defining the lower and upper bound of an
$n$-dimensional cube of side length $1$.  For the five-point stencil
example there are five unit spans.
Figure~\ref{fig:inference-steps-informal}(a) illustrates these.

The next step repeatedly coalesces any contiguous spans until a fixed
point is reached, resulting in a set of (possibly overlapping) spans
covering the original space of schemes. Two $n$-dimensional spans
$\vect{x}$ and $\vect{y}$ are contiguous if for all but one
$j \in \{1, \ldots, n\}$ then $\vect{x}_j = \vect{y}_j$ and for one
$j$ then $\vect{x}^U_1 + 1 = \vect{y}^U$. For example, spans
$\left[\vtwoh{-1}{0}\; \vtwoh{-1}{0}\right]$ and
$[\vtwoh{0}{0} \; \vtwoh{1}{0}]$ are contiguous in the first dimension
and can be coalesced into the span $[\vtwoh{-1}{0} \; \vtwoh{1}{0}]$.

\begin{definition}\label{def:span-coalesc}
  Span coalescing $\bullet$ is defined for spans contiguous in the
  first dimension:
%%
\begin{align*}
\vect{x} \bullet \vect{y}
= &
\begin{cases}
(\vect{x}^L, \vect{y}^U) & \vect{x}^U_1 + 1 = \vect{y}^L_1 \; \wedge \;
\vect{x}_{2:n} = \vect{y}_{2:n}) \\
\bot  & \textit{otherwise}
\end{cases}
% $m$-dimensions into coalesced spans which vary in
% $m+1$-dimensions.
\end{align*}
%%
\end{definition}
\noindent
To coalesce regions which are contiguous in other dimensions
other than the first, we generate all rotations of the spans in
$n$-dimensional space by permutation on spans, and fold together contiguous
regions using $\bullet$ (above). This
generates all possible ways of coalescing regions. For our example,
the unit spans in Figure~\ref{fig:inference-steps-informal}(a)
are coalesced into the contiguous spans in
Figure~\ref{fig:inference-steps-informal}(b).
This procedure is iterated until a fixed point is reached. In this
example this is reached in the first step.
 Then any spans that are completely contained by any other span are deleted,
leaving a minimal set of spans (which may overlap, but none of which
fully contains another), Figure~\ref{fig:inference-steps-informal}(c).
The final spans are used to synthesise concrete stencil specification
syntax (\S~\ref{sec:inf-step4}), where
each span becomes a product \texttt{*} of regions which are
combined by the sum \texttt{+}.

More formally the algorithm proceeds as follows.
For all $a \in \mathsf{dom}(U)$, let $M = U(a)$. Every scheme in $M$
is mapped to a unit span
by $\textsf{unitSpans}(M) = \{[u \; u] \mid u\!\in\!M\}$. For our example:
%%
\begin{align*}
  N(U(\texttt{a})) = \stwo{0}{0}{0}{0} \stwo{1}{0}{1}{0} \stwo{-1}{0}{-1}{0} \stwo{0}{1}{0}{1} \stwo{0}{-1}{0}{-1}
\end{align*}
%%
In the following we let $M_0 = N(U(\texttt{a}))$ for our example.

Each set of unit spans is applied to the fixed point of the \spanOp{} function
(defined shortly), which coalesces spans into contiguous regions. That is, we compute
 $(\mu \, \spanOp)$ on input $N(M)$, where \spanOp{} is defined by the steps:

\begin{figure}[t]
\vspace{-0.8em}
\hspace{-0.5em}
\begin{minipage}{0.16\linewidth}
\begin{align*}
\setlength{\arraycolsep}{0em}
\begin{array}{c}
\quad \\[0em]
\begin{tikzpicture}[scale=0.8]
%\draw[step=0.5cm,lightgray,very thin] (0.5,0.5) grid (2,2);
\node at (1.25,1.22) {\textnormal{\footnotesize{0,0}}};
\draw[thick] (1,1) rectangle (1.5,1.5);
\draw[thick] (1.5,1) rectangle (2,1.5);
\draw[thick] (1,0.5) rectangle (1.5,1);
\draw[thick] (0.5,1) rectangle (1,1.5);
\draw[thick] (1,1.5) rectangle (1.5,2);
\end{tikzpicture} \\[0em]
(a)
\end{array}
\end{align*}
\end{minipage} $\rightarrow$ \begin{minipage}{0.43\linewidth}
\begin{align*}
\setlength{\arraycolsep}{0em}
\begin{array}{c}
\begin{tikzpicture}[scale=0.72]
%\draw[step=0.5cm,lightgray,very thin] (0.5,0.5) grid (2,2);
%\node at (1.25,1.22) {\textnormal{\small{0,0}}};
\draw [black, thin] (0.35,0) to [round left paren ] (0.35,2);
\draw[thick] (0.5,1) rectangle (2,1.5);
\draw[thick] (1,0.40) rectangle (1.5,0.90);
\draw[thick] (1,1.60) rectangle (1.5,2.10);
\node at (2.3,1.22) {+};
\draw[thick] (3.8,1) rectangle (4.3,1.5);
\draw[thick] (2.6,1) rectangle (3.1,1.5);
\draw[thick] (3.2,0.5) rectangle (3.7,2);
\draw [black, thin] (4.4,0) to [round right paren ] (4.4,2);
\end{tikzpicture} \\[-0.5em] (b)
\end{array}
\end{align*}
\end{minipage}$\rightarrow$ \begin{minipage}{0.25\linewidth}
\begin{align*}
\setlength{\arraycolsep}{0em}
\begin{array}{c}
\begin{tikzpicture}[scale=0.72]
%\node at (1.25,1.22) {\textnormal{\small{0,0}}};
\draw [black, thin] (0.8,0) to [round left paren ] (0.8,2);
\draw[thick] (1,0.5) rectangle (1.5,2);
\node at (1.8,1.22) {+};
\draw[thick] (2.1,1) rectangle (3.6,1.5);
\draw [black, thin] (3.7,0) to [round right paren ] (3.7,2);
%\node at (2.75,1.22) {\textnormal{\small{0,0}}};
\end{tikzpicture} \\[-0.5em] (c)
\end{array}
\end{align*}
\end{minipage}
\caption{Illustration of grouping spans in
specification inference}
\label{fig:inference-steps-informal}
\vspace{-0.8em}
\end{figure}

\noindent
\textbf{1)\,} Compute all permutations on the column vectors in a span, \ie{},
  $[\vect{x}^L \vect{x}^U] \mapsto [\pi\vect{x}^L \, \pi\vect{x}^U]$
for a permutation $\pi$. For each permutation function $\pi^n_i$
(the $i$-th permutation for vectors of size $n$) we pair the
permutation function with the set of permuted spans so that
the spans can be un-permuted later.
%
\begin{equation*}
P(M) = \bigcup_{i \in n!} (\pi^n_{i} , \; \{[\pi^n_i
\vect{x}^L \; \pi^n_i\vect{x}^U] \, \mid \, [\vect{x}^L \; \vect{x}^U]
\leftarrow M\}
\end{equation*}
%
This corresponds to all rotations of the space.  For our example:
%
\begin{align*}
P(M_0) =
\{&(\pi^2_1, \{ \stwo{0}{0}{0}{0},
\stwo{1}{0}{1}{0},
\stwo{-1}{0}{-1}{0},
\stwo{0}{1}{0}{1},
\stwo{0}{-1}{0}{-1} \})
\\
&(\pi^2_2, \{
 \stwo{0}{0}{0}{0},
 \stwo{0}{1}{0}{1},
 \stwo{0}{-1}{0}{-1},
 \stwo{1}{0}{1}{0},
 \stwo{-1}{0}{-1}{0}\})\}
\end{align*}
%
where $\pi^2_1$ is the identity permutation and $\pi^2_2$ is the
permutation flips the order of the two elements in the
vectors.

\noindent
\textbf{2)\,} Sort each permutation set into an ordered list, by the ordering:
\begin{equation*}
  \vect{x} \leq \vect{y} =
      \exists i \, . \; \vect{x}^L_{i} \leq \vect{y}^L_{i} \; \wedge \;
        (i = n \vee \vect{x}_{i+1:n} = \vect{y}_{i+1:n})
\end{equation*}
%
giving the sorting function $\textsf{sort}(M)$. For our example:
%%
\begin{align*}
\textsf{sort}(P(M_0)) =
\{&(\pi^2_1, [
\stwo{0}{-1}{0}{-1},
\stwo{-1}{0}{-1}{0},
\stwo{0}{0}{0}{0},
\stwo{1}{0}{1}{0},
\stwo{0}{1}{0}{1}] )
\\
&(\pi^2_2, [
\stwo{0}{-1}{0}{-1},
\stwo{-1}{0}{-1}{0},
\stwo{0}{0}{0}{0},
\stwo{1}{0}{1}{0},
\stwo{0}{1}{0}{1}])\}
\end{align*}
\noindent
\textbf{3)\,} Reduce each list pairwise by $\bullet$ operation
(Defn.~\ref{def:span-coalesc}) to coalesce contiguous regions,
given by $\textsf{fold}\bullet$. For our example, this yields:
%
\begin{align*}
\{&(\pi^2_1, [
\stwo{0}{-1}{0}{-1},
\stwo{-1}{0}{1}{0},
\stwo{0}{1}{0}{1}
]), (\pi^2_2, [
\stwo{0}{-1}{0}{-1},
\stwo{-1}{0}{1}{0},
\stwo{0}{1}{0}{1}])\}
\end{align*}
%
\textbf{4)\,} Un-permute and union together, \ie{},
%
\[
U(M) = \bigcup \{[\pi \vect{x}^L, \pi \vect{x}^U]
 \mid [\vect{x}^L, \vect{x}^U] \leftarrow S, (\pi, S) \leftarrow M\}
\]
For our example:
%
\begin{align*}
U(M) =
\{\stwo{0}{-1}{0}{-1},
\stwo{-1}{0}{1}{0},
\stwo{0}{1}{0}{1},
\stwo{-1}{0}{-1}{0},
\stwo{0}{-1}{0}{1},
\stwo{1}{0}{1}{0}\}
\end{align*}
%
\textbf{5)\,} Filter by the region containment predicate $\containedin$, that
  is, if any region is contained within another then remove the
  smaller, defined:
%
\begin{align*}
& \vect{x} \containedin \vect{y} = \vect{y}^L_1 \leq \vect{x}^L_1 \wedge \vect{x}^U_1 \leq \vect{y}^U_1
  \wedge (\vect{x}^L_{2:n}, \vect{x}^U_{2:n}) \containedin
  (\vect{y}^L_{2:n}, \vect{y}^U_{2:n})
\end{align*}
For our example this then yields the final result of:
\begin{align*}
\spanOp(M)
= \{\stwo{-1}{0}{1}{0} \stwo{0}{-1}{0}{1}\}
\end{align*}
since $\stwo{0}{1}{0}{1} \sqsubseteq \stwo{0}{-1}{0}{1}$
and $\stwo{1}{0}{1}{0} \sqsubseteq \stwo{-1}{0}{1}{0}$.
Applying \spanOp{} again yields the same
result for our example; the fixed point is reached.

\subsubsection{Step 4: Synthesise specifications from spans}
\label{sec:inf-step4}

\newcommand{\oplusbig}{\operatornamewithlimits{\term{\Large{$\oplus$}}}}
\newcommand{\odobig}{\operatornamewithlimits{\term{\Large{$\odot$}}}}
\newcommand{\bplus}{\operatornamewithlimits{\term{\Large{+}}}}
\newcommand{\tySum}[1]{#1^{\term{+}}}
\newcommand{\tyProd}[1]{#1^{\term{*}}}
\newcommand{\specDNF}{\textbf{spec}}

Abstract syntax trees of the specification language
are synthesised from the set $\spanOp(M)$ of coverings spans.
This conversion is defined in terms
of an algebra on \textit{spec} syntax in disjunctive-normal form
(which we denote \specDNF{}) and which is wrapped
in the $\textsf{Approx}$ type (see
Defn.~\ref{def:mult-and-approx},
p.~\pageref{def:mult-and-approx}).  The algebra mirrors the shape
of the region combinators for sum and product,
with operators $\oplus$ and $\odot$ and a constructor for \specDNF{}:
%%
\begin{align*}
\oplus & : \mathsf{Approx}(\specDNF) \rightarrow \mathsf{Approx}(\specDNF) \rightarrow
  \mathsf{Approx}(\specDNF) \\
\odot & : \mathsf{Approx}(\specDNF) \rightarrow \mathsf{Approx}(\specDNF) \rightarrow
        \mathsf{Approx}(\specDNF) \\
\mathsf{toSpec} & : \mathbb{N}_{>0} \rightarrow (\domainVal{} \times \domainVal{}) \rightarrow \mathsf{Approx}(\specDNF)
\end{align*}
%%
The $\mathsf{toSpec}$ function maps a dimension identifier
and a pair of lower and upper bound
values (one-dimensional) to a specification with a one-dimensional region.
An entire span is converted into an $n$-dimensional
specification by applying $\mathsf{toSpec}$ to each
pair of lower and upper bounds in a span, per dimension, and multiplying these
specifications by $\odot$. The resulting specifications for
each span are then summed by $\oplus$:
\begin{equation*}
\textsf{spec}(M) =
\oplusbig\limits_{\vect{x} \in \spanOp(M)} \,
\odobig\limits_{i \in \{1, \ldots, n\}} \, \mathsf{toSpec} \; i \; (\vect{x}^L_i, \vect{x}^U_i)
\end{equation*}
We unpack the definitions of this algebra.

\begin{definition}[Specifications in DNF]
Let $r_1, ..., r_n$ range over elements of the \textit{region}
syntax (see Figure~\ref{fig:syntax}, p.~\pageref{fig:syntax}).
The set \specDNF{} is a subset of syntax trees \textit{spec},
containing specifications in disjunctive normal form over regions
where the disjunct is \term{+} (since its model is in terms of set
union) and the conjunct is \term{*} (since its model is a form of tensor).
That is elements of \specDNF{} are of the form:
%%
$
(r^1_{1} \; \term{*} \ldots \term{*} \; r^1_n)\; \term{+} \; \ldots
\term{+} \; (r^m_1 \, \term{*} \; \ldots \term{*} \; r^m_p)
$.

More formally, $\specDNF \subseteq (\textit{spec} \cup \{\epsilon\})$, defined:
%%
\begin{align*}
\specDNF{} & ::= \tySum{\specDNF} \mid \epsilon \\
\tySum{\specDNF} & ::= \tyProd{\specDNF} \; \term{+} \; \tySum{\specDNF} \mid
  \tyProd{\specDNF} \\
\tyProd{\specDNF} & ::= \textit{region} \; \term{*} \; \tyProd{\specDNF} \mid
   \textit{region}
\end{align*}
%%
We include the empty element $\epsilon$ which is not present in the
syntax.
In the following, $S, T$ range over elements of
\specDNF{} and $\tyProd{s}, \tyProd{t}$ range over elements
of $\tyProd{\specDNF}$ (for example, $\tyProd{s} = r^1_{1} \; \term{*}
\ldots \term{*} \; r^1_n$).
\end{definition}

\newcommand{\exactR}{\mathsf{exact}\;}
\newcommand{\exact}[1]{\exactR(#1)}
\newcommand{\upperB}[1]{\mathsf{upper}\;(#1)}

\begin{figure}[t]
%\newcommand{\llower}{\vect{x}^L_i}
%\newcommand{\lupper}{\vect{x}^U_i}
\vspace{-0.7em}
\newcommand{\llower}{l}
\newcommand{\lupper}{u}
{\scalebox{0.97}{
\begin{minipage}{1\linewidth}
\begin{align*}
&\mathsf{toSpec} \; d \; (\llower, \lupper) = \\
&\hspace{-0.4em}\begin{cases}
%% POINT
\exact{\stenReflS{d}} & \llower = 0 \, \wedge \, \lupper = 0\\[0.2em]
%% FWD
\exact{\stenFwdS{\lupper}{d}}  & \llower = 0 \, \wedge \, \lupper > 0 \\[0.2em]
\exact{\stenFwdSR{\lupper}{d}{\irreflS}}\!\!\!\! & \llower = 1 \, \wedge \, \lupper > 0 \\[0.2em]
%% BWD
\exact{\stenBwdS{|\llower|}{d}}  & \llower < 0 \, \wedge \, \lupper = 0 \\[0.2em]
\exact{\stenBwdSR{|\llower|}{d}{\irreflS}}\!\!\!\! & \llower < 0 \, \wedge \, \lupper = -1 \\[0.2em]
%% CEN
\exact{\stenCenS{\lupper}{d}}& \llower < 0 \wedge \lupper > 0 \wedge |\llower| = \lupper \\[0.2em]
\exactR(\stenBwdS{|\llower|}{d} & \llower < 0 \wedge \lupper > 0  \wedge |\llower| \neq \lupper \\
\qquad\qquad \texttt{+} \, \stenFwdS{\lupper}{d}) & \\[0.2em]
% BOUND
\upperB{\stenFwdS{\lupper}{d}} & \llower > 1 \\[0.2em]
\upperB{\stenBwdS{|\llower|}{d}} \quad & \llower < (-1)
 \\[0.2em]
% ABSOLUTE REP
\exact{\epsilon} & \llower = \infty \, \vee \, \lupper = \infty
  \end{cases}
\end{align*}
\end{minipage}
}}
\caption{\textsf{toSpec} constructor of $\mathsf{Approx}(\specDNF)$ values.}
\label{fig:to-spec}
\vspace{-0.8em}
\end{figure}

\begin{definition} The $\mathsf{toSpec}$ constructor is defined
in Figure~\ref{fig:to-spec}.
The first two cases map to \texttt{forward} stencils since the lower
bound is $0$ or $1$, with the depth determined by the upper bound of
the span. If the lower bound is $1$, then the
$0$-offset point is missing so the region is given a
\texttt{nonpointed} attribute. The second two cases are similar for
\texttt{backward} but now the depth of the region is determined by the
absolute of the lower bound. In the fifth-case, where the lower bound
is negative and the upper bound is positive and they have the same
magnitude, then a \texttt{centered} region is constructed with
the depth as the common magnitude. If the magnitudes are not the same
(sixth case) then separated \texttt{forward} and \texttt{backward}
regions are summed, with their corresponding depths.

If the lower bound is greater than 1
(which implies the upper bound is greater than 1 too) then
this represents a region that is not at the origin (or next to) the
origin and thus cannot be represented in the specification language.
Thus, an upper bound is instead returned showing the maximum extent
of the region. The penultimate case is the dual, when the lower bound is less
than $-1$, giving an upper bound on a \texttt{backward} specification.

Finally, if either bound is $\infty$, representing a non
neighbourhood index, then an empty specification is given
 with no regions, indicating that there is not even a
pointed 0-offset access in this dimension.

Note, $\mathsf{toSpec} \, d \, (l, u)$ is always defined
 if $l$ less than or equal to $u$.
\end{definition}

\begin{definition} The $\odot$ operation on $\textsf{Approx}(\specDNF{})$
  is defined in terms of the intermediate operator $\hat{\odot}$ on
  \specDNF{}, defined:
%%
\begin{align*}
S \,\hat{\odot}\, \epsilon = S \quad
\epsilon \,\hat{\odot}\, S = S \qquad
S \,\hat{\odot}\, T = \bplus_{(\tyProd{s}, \tyProd{t}) \in (S \times
   T)} (\tyProd{s} \; \term{*} \; \tyProd{t})
\end{align*}
%%
That is, the $\hat{\odot}$ product of two specifications $S$ and $T$ in DNF form
is the $\term{+}$ sum of all pairwise $\term{*}$ products for every
pair of $\tyProd{spec}$ drawn from $S$ and $T$. The result
is indeed in DNF-form, where $\hat{\odot}$ is essentially applying
 the (\trule{\textsc{dist}}) rule of our equational theory
 (Fig.~\ref{fig:equations},p.~\pageref{fig:equations}). The operation is
commutative and associative, and sound with respect to our
model (see Lemma~\ref{lem:alg-soundness} below).
The $\hat{\odot}$ operation is then lifted to all combinations of
\textsf{Approx} for $\odot$:
\begin{align*}
\textsf{lower}(S) \odot \textsf{exact}(T) = \; &
       %\textsf{exact}(T) \odot \textsf{lower}(S) = \;
       \textsf{both} \, (S \hat{\odot} T, \; T) \\
\textsf{upper}(S) \odot \textsf{exact}(T) = \; &
       %\textsf{exact}(T) \odot \textsf{upper}(S) =
       \textsf{both} \, (T, \; S \hat{\odot} T) \\
\textsf{lower}(S) \odot \textsf{both}(T_l,T_u) = \; &
        %\textsf{both}(T_l,T_u) \odot \textsf{lower}(S) = \;
        \textsf{both} \, (S \hat{\odot} T_l, \; T_u) \\
\textsf{upper}(S) \odot \textsf{both}(T_l,T_u) = \; & \textsf{both} \,
                                                      (T_l, \; S \hat{\odot} T_u) \\
\textsf{both}(S_l,S_u) \odot
\textsf{both}(T_l,T_u)  = \; & \textsf{both} (S_l \hat{\odot} T_l, \; S_u
                         \hat{\odot} T_u) \\
\textit{inj}(S) \odot \textit{inj}(T) = \; & \textit{inj}(S
                                              \hat{\odot} T)
\end{align*}
where in the last case $\textit{inj}$ corresponds to the unary
injections of \textsf{Approx}. For brevity we omit the cases where the arguments above are flipped
as $\odot$ is commutative.
\end{definition}

\begin{definition}The $\oplus$ operation on $\textsf{Approx}(\specDNF)$
is defined in a similar way to $\odot$, with the intermediate $\hat{\oplus}$ on
$\specDNF$ given by:
%%
\begin{equation*}
S \,\hat{\oplus}\, \epsilon = S \quad
\epsilon \,\hat{\oplus}\, S = S \quad
S \, \hat{\oplus} \, T = S \, \term{+} \, T
\end{equation*}
%%
Thus, if neither specification is empty, their $\hat{\oplus}$ is just
the syntactic sum $\term{+}$. Then $\oplus$ is the lifting of
$\hat{\oplus}$ to \textsf{Approx} with the shape as $\odot$ in the previous
definition.
\end{definition}

\begin{lemma}[$\hat{\odot}$, $\hat{\oplus}$ soundness]$\forall S,
  T \in \specDNF$ where $S\!\neq \epsilon \wedge T\!\neq \epsilon$:
\begin{align*}
\interp{S \; \hat{\odot} \; T}_n = \interp{S}_n \otimes \interp{T}_n
  \qquad
\interp{S \; \hat{\oplus} \; T}_n = \interp{S}_n \cup \interp{T}_n
\end{align*}
\label{lem:alg-soundness}
\vspace{-2em}
\end{lemma}
%
\noindent
Thus our model validates the specification synthesis process which concludes the
inference procedure. For our example, we infer and then synthesise the
5-point specification:
%%
\begin{minted}[linenos=false]{fortran}
!= stencil centered(depth=1, dim=1)*pointed(dim=2) + centered(depth=1, dim=2)*pointed(dim=1):: a
\end{minted}
%%

\section{Evaluation}

We have processed nearly one million lines of Fortran code using our
CamFort tool and examined the output to try and determine how useful
and applicable it may be. The body of Fortran code is collected from
many computational science projects for which we believe our tool will
be potentially helpful.

% Corpus details
% specfem3d ``simulates acoustic (fluid), elastic (solid), coupled
% acoustic/elastic, poroelastic or seismic wave propagation in any
% type of conforming mesh of hexahedra (structured or not).''
% https://github.com/geodynamics/specfem3d

% arpack-ng - Fortran 77 routines designed to solve large scaled
% eigenvalue problems - https://github.com/opencollab/arpack-ng

\subsection{Core hypotheses}

\paragraph{Hypothesis 1: Stencil computations
(as in Definition~\ref{def:stencil}) are a common programming pattern
in numerical code}

%    numStencilSpecs: 60525
% 4338 files, linesTotal: 1357377, linesParsed: 959427
% Number of stencils (tickAssign) / number of loops
% Percentage of a code base that is taken up by stencils

Out of 1.35 million lines of code that we fed into our program, and
the 959,427 lines that we were able to parse, we identified 60,525
cases that fit our criteria for stencil specification.

\paragraph{Hypothesis 2: Most stencil computations
have a regular pattern which can be captured with a simple
language of contiguous regions}

% Number of stencils we give a spec to (where a spec declaration to n
% variables is counted n times) as percentage of number of stencils

    % tickAssign: 97439
    % tickAssignSuccess: 29402 % <- number of times tickAssign is accompanied by a successful stencil inference
                               % counts only by +1 each time that tickAssign finds a stencil

In the analysed body of code we found that we were able to deduce
some form of stencil specification from approximately 30\% of the
potential stencils examined.

%
\subsection{Analysing the design of our specifications}

% 1. How may are just reflexive and how many some reflexive
% justReflexive: 55504
% someReflexive: 15823

55,504 specifications involved only reflexive regions, of which 15,823
only had specifications for some of the dimensions, while 39,681 were
reflexive on all dimensions.

% (/ 55504.0 60525) = 91.7%
% (/ 15823.0 60525) = 26.1%
% 91.7\% of the specifications involved only reflexive regions. 26.1\%

% 2. Break down of "interesting specifications" (contain more than
% just reflexive)
%   - Single-action, histogram on depths
%   - Single-action + irreflexives, histogram on depths
%   - Bounded specs (how many lower, upper, or both)

Bounded specifications were fairly unusual, with only 157
\texttt{atMost} and 36 \texttt{atLeast}, the latter of which were
always also bounded on both sides.
    % atLeast: 36
    % atMost: 157
    % boundedBoth: 36

%   - Multi-action combined only by *
%   - Multi-action combined also by +, histogram of tree depths

% Present not just percentages but concrete numbers

\section{Related work}\label{sec:related-work}

\citet{kamil2016verified} propose verified lifting to extract a mathematical
representation of low-level, and potentially optimised, stencils in Fortran
code. Their aim is to use the extracted predicate to pipe to the \textsf{Halide}~\citep{ragan2013halide} high-performance
compiler and so they need to capture the full meaning of the stencil computation.

% ACR-I think we could skip this in the interests of space
%This led them to develop \textsf{STNG}, a loop invariant and postcondition
%finder through syntax-guided synthesis. Using this approach they restrict the
%search space of postconditions to predicates of a certain form, one that is
%compatible with \textsf{Halide}. They then look at concrete loop iterations and
%form an hypothesis about the invariant, which is later confirmed or put back
%into the system using a SMT solver.

CamFort differs from this significantly. The aim of CamFort is not a
full representation of the stencil, but enough information to capture the access
pattern. Hence, we do not require SMT solving or search of loop invariants. It
suffices to do comparatively simple data-flow analysis.

Due to the sheer amount of analysis involved, Kamil et al. report that some
degenerate kernels take up to 17 hours to analyse and others require programmer
intervention for correct invariants to be inferred. Since CamFort mostly does
syntactic analysis, the inference procedure terminates quickly and never requires programmer intervention. The downside of our
approach is that if optimisation causes the stencil computation to be heavily
obfuscated, \textsf{STNG} would capture the access behaviour better.

Most fundamentally, unlike CamFort, Kamil et al. does not provide a syntactic
representation of their postcondition, so there is no way of future-proofing the
code for future edits. Even if they were to provide a syntactic representation,
for complex stencils such as Navier-Stokes from \S~\ref{sec:intro}, it would be
as verbose as the code itself making it difficult for programmer to understand
the overall shape of the indexing behaviour.

\section{Conclusion}\label{sec:conclusion}

\bibliography{references}

\end{document}

%%  LocalWords:  refactoring affine parameterised nonpointed atMost
%%  LocalWords:  centered readOnce atLeast discretisation Equational
%%  LocalWords:  equational disjunction denotational dimensionality
%%  LocalWords:  interprocedural Fortran CamFort preprocessor
%%  LocalWords:  committers