### 2017-03-11 array17.tex
Array 17 first draft

\documentclass[10pt,preprint,numbers]{sigplanconf}

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{stmaryrd}
\usepackage{color}
\usepackage{graphics}
\usepackage{fancyvrb}
\usepackage{subfigure}
\usepackage{multirow}
\usepackage{minted}
\usepackage{breqn}
\usepackage{enumitem}
%\usepackage{subcaption}
\usepackage{siunitx} % For pretty-printing numeric values and SI units
                     % of measure. e.g., the tabular column type S is
                     % used to print nice-looking tables of numbers.
\sisetup{ % defaults
  group-separator={,},
  group-minimum-digits={3},
  output-decimal-marker={.},
  table-format = 6
}

\usepackage{tikz}
\usepackage{xypic}
\usepackage{natbib}

\usetikzlibrary{calc}

%% http://tex.stackexchange.com/questions/55068/is-there-a-tikz-equivalent-to-the-pstricks-ncbar-command
\tikzset{
    ncbar angle/.initial=90,
    ncbar/.style={
        to path=(\tikztostart)
        -- ($(\tikztostart)!#1!\pgfkeysvalueof{/tikz/ncbar angle}:(\tikztotarget)$)
        -- ($(\tikztotarget)!($(\tikztostart)!#1!\pgfkeysvalueof{/tikz/ncbar angle}:(\tikztotarget)$)!\pgfkeysvalueof{/tikz/ncbar angle}:(\tikztostart)$)
        -- (\tikztotarget)
    },
    ncbar/.default=0.5cm,
}

\tikzset{round left paren/.style={ncbar=0.4cm,out=110,in=-110}}
\tikzset{round right paren/.style={ncbar=0.4cm,out=70,in=-70}}

\fvset{
  linenos=true,
  fontsize=\footnotesize,
  breaklines=true,
 breakafter=w),
  xleftmargin=\parindent,
  numbersep=1em
}

\usemintedstyle{vs}

\definecolor{darkblue}{rgb}{0.0,0.0,0.5}
\definecolor{darkgreen}{rgb}{0.0,0.4,0.0}
\definecolor{darkdarkgreen}{rgb}{0.0,0.35,0.0}

\usepackage{hyperref}
\hypersetup{
    unicode=false,          % non-Latin characters in Acrobat's bookmarks
    pdftoolbar=true,        % show Acrobat toolbar?
    pdfmenubar=true,        % show Acrobat menu?
    pdffitwindow=false,      % page fit to window when opened
    pdftitle={},    % title
    pdfauthor={}
    pdfsubject={},   % subject of the document
    pdfnewwindow=true,      % links in new window
    pdfkeywords={keywords}, % list of keywords
    colorlinks=true,       % false: boxed links; true: colored links
    linkcolor=darkblue,          % color of internal links
    citecolor=darkblue,        % color of links to bibliography
    filecolor=green,      % color of file links
    urlcolor=blue,          % color of external links
}

\usepackage{cleveref}
\crefname{section}{\S}{\S\S}
\Crefname{section}{\S}{\S\S}
\crefname{subsection}{\S}{\S\S\S}
\Crefname{subsection}{\S}{\S\S\S}
\crefname{figure}{fig.}{}
\Crefname{figure}{Fig.}{}
\crefname{definition}{def.}{}
\Crefname{definition}{Def.}{}

\CustomVerbatimEnvironment{SpecVerbatim}{Verbatim}{fontsize=\footnotesize,xleftmargin=0.5cm,
xrightmargin=0.2cm,commandchars=\\\{\},baselinestretch=0.98,numbersep=0.9em}
\CustomVerbatimEnvironment{ExmVerbatim}{Verbatim}{fontsize=\footnotesize,xleftmargin=0.5cm,
xrightmargin=0.2cm,baselinestretch=0.98,numbers=left,numbersep=0.9em,commandchars=\\\{\}}
\CustomVerbatimEnvironment{IVerbatim}{Verbatim}{fontsize=\relsize{-1},xleftmargin=0.5cm,
xrightmargin=0.2cm,commandchars=\\\{\},baselinestretch=0.98,numbersep=0.9em}

\definecolor{darkgreen}{rgb}{0.0,0.5,0.0}
\definecolor{darkpurple}{rgb}{0.6,0.0,0.6}
\definecolor{orange}{rgb}{0.8,0.4,0.0}
\definecolor{darkorange}{rgb}{0.5,0.2,0.0}
\definecolor{marco}{rgb}{0.0,0.3,0.5}
\definecolor{gray}{rgb}{0.2,0.2,0.2}

\newcommand\mdef{\stackrel{\mathclap{\normalfont\mbox{def}}}{=}}

\newcommand{\bn}{\mathbb{N}}

\newcommand{\todo}[1]{\textcolor{blue}{Dom: #1}}
\newcommand{\dnote}[1]{\textcolor{darkpurple}{Dom: #1}}
\newcommand{\mnote}[1]{\textcolor{darkgreen}{Mistral: #1}}
\newcommand{\anote}[1]{\textcolor{red}{Andy: #1}}

\newcounter{block}

\newtheorem{lemma}[block]{Lemma}
\newtheorem{proposition}[block]{Proposition}

\theoremstyle{definition}

\newtheorem{theorem}[block]{Theorem}
\newtheorem{remark}[block]{Remark}
\newtheorem{example}[block]{Example}
\newtheorem{definition}[block]{Definition}

% Writing macros
\newcommand{\ie}{\emph{i.e.}}
\newcommand{\eg}{\emph{e.g.}}

\newcommand{\dimId}{\texttt{dim}}

% Semantics related
\newcommand{\interp}[1]{\llbracket{#1}\rrbracket}

% Syntax macros
\newcommand{\nonterm}[1]{\textit{#1}}
\newcommand{\term}[1]{\texttt{#1}}

\newcommand{\stenRefl}[1]{\term{pointed} \, (\term{dim=}#1)}
\newcommand{\stenFwd}[3]{\term{forward} \, (\term{depth=}#1,
  \term{dim=}#2{#3})}
\newcommand{\stenBwd}[3]{\term{backward} \, (\term{depth=}#1,
  \term{dim=}#2{#3})}
\newcommand{\stenCen}[3]{\term{centered} \, (\term{depth=}#1,
  \term{dim=}#2{#3})}
\newcommand{\irrefl}{\texttt{nonpointed}}

\newcommand{\stenReflS}[1]{\term{pointed} \, (\term{dim=}#1)}
\newcommand{\stenFwdS}[2]{\term{fwd} \, (\term{depth=}#1,
  \term{dim=}#2)}
\newcommand{\stenBwdS}[2]{\term{bwd} \, (\term{depth=}#1,
  \term{dim=}#2)}
\newcommand{\stenCenS}[2]{\term{cen} \, (\term{depth=}#1,
  \term{dim=}#2)}
\newcommand{\irreflS}{\texttt{np}}

\newcommand{\stenFwdSR}[3]{\term{fwd} (\term{depth=}#1,
  \term{dim=}#2, #3)}
\newcommand{\stenBwdSR}[3]{\term{bwd} (\term{depth=}#1,
  \term{dim=}#2, #3)}
\newcommand{\stenCenSR}[3]{\term{cen} (\term{depth=}#1,
  \term{dim=}#2, #3)}
\newcommand{\stenReflSR}[1]{\term{point} (\term{dim=}#1)}

% SYNTAX OPERATIONS AND PREDICATES
%\newcommand{\neigh}{\textsf{neigh}}
\newcommand{\arrayTy}{\textsf{array}}
\newcommand{\rhsExp}{\textsf{rhsExp}}
\newcommand{\var}{\textsf{var}}

%% VECTOR NOTATIONS
\newcommand{\vect}[1]{\textbf{#1}}

\newcommand{\vtwohs}[2]{\setlength{\arraycolsep}{0em}
\left[\begin{array}{cc}#1 \, & \, #2\end{array}\right]\!}

\newcommand{\vtwoh}[2]{\setlength{\arraycolsep}{0em}
\left[\begin{array}{cc}#1 \; & \; #2\end{array}\right]}

\newcommand{\vthreeh}[3]{\setlength{\arraycolsep}{0em}
\left[\begin{array}{ccc}#1 \; & \; #2 \; & \; #3\end{array}\right]}
\newcommand{\vtwo}[2]{\setlength{\arraycolsep}{0em}
\left[\begin{array}{l}#1\\#2\end{array}\right]}
\newcommand{\vthree}[3]{\setlength{\arraycolsep}{0em}
\left[\begin{array}{l}$#1$\\$#2$\\$#3$\end{array}\right]}
\newcommand{\stwo}[4]
%{\vtwo{#1}{#2}\!\vtwo{#3}{#4}}
{\setlength{\arraycolsep}{0.1em}
\left[\begin{array}{rr}$#1$ & $#3$\\$#2$ & $#4$\end{array}\right]}

\newcommand{\singleEntry}[2]{\textbf{J}_{#2}^{#1}}
\newcommand{\zeroEntry}[2]{\textbf{K}_{#2}^{#1}}

%% OPERATIONS ON SPANS and VECTORS
\newcommand{\containedin}{\sqsubseteq}

%% MODEL
\newcommand{\effdims}[2]{\mathit{constr}(#1)_{#2}}

\newcommand{\trule}[1]{{\footnotesize{(\text{#1})}}}
%\include{results}

%\newcommand{\cons}[3]{\mathbf{C}_{#3}(#1, #2)}
\newcommand{\consName}{\textbf{agree\textprime}}
\newcommand{\cons}[2]{#1 \,\, \consName{} \,\, #2}
\newcommand{\consAName}{\textbf{agree}}
\newcommand{\consA}[2]{#1 \,\, \consAName{} \,\, #2}
\newcommand{\consSub}[3]{#2 \,\, \consAName{}_{#1} \,\,#3}

\title{Verifying Spatial Properties of Stencil Computations using CamFort}
%\titlebanner{DRAFT -- Do not distribute}
\authorinfo{}{}{}
%\preprintfooter{DO NOT DISTRIBUTE}

\begin{document}
\maketitle

\bibliographystyle{plainnat}

\begin{abstract}
  Stencil computations are a common programming pattern in numerical code and
  typically form the core of graphics and scientific computing applications.
  However, they are often complex and unwieldy, involving substantial
  fine-grained manipulation of array indices across multiple arrays and
  dimensions. This requires careful testing, which is almost always done
  manually and then discarded. In this paper, we propose an alternate approach:
  a specification language and automated tool for verifying the spatial
  behaviour of stencils. The specification language is based on the hypothesis
  that stencil computations read from arrays with a regular, fixed pattern
  which can be captured by a simple set of abstract combinators. These
  specifications can be embedded as annotations in source code, against which
  code can be checked for conformance. For legacy code, spatial specifications
  can be inferred, providing documentation and aiding future software
  maintenance.

  We evaluate our specification language and verification tool
  against a corpus of numerical Fortran code ($\sim$ 1 million lines) for which
  we generate $\sim$ 60,000 specifications, showing the vast majority of stencil
  computations indeed have a simple, regular, static shape.
\end{abstract}

%\keywords{program verification, specification, stencils}

\section{Introduction}\label{sec:intro}

\emph{Stencils} are a ubiquitous programming pattern, common in scientific and
numerical computing applications. Informally, a stencil computation computes an
array whose value at each index $i$ is calculated from a \emph{neighbourhood}
of values around $i$ in some input array(s), \eg{}, the Game of Life,
convolutions in image processing, approximations to differential equations. For
example, the following computes the one-dimensional discrete Laplace transform
(an approximation to a derivative) in Fortran:
%
\begin{minted}{fortran}
do i = 1, (n-1)
      b(i) = a(i-1) - 2*a(i) + a(i+1)
end do
\end{minted}
%
Values \texttt{b(i)} are calculated from a neighbourhood of elements around
\texttt{i} in the input array \texttt{a}. In this example, the access pattern
is simple and easily understood. More complex stencil computations are more
prone to errors from simple lexical mistakes. For example, the following is a
small snippet from a Navier-Stokes fluid simulator (based
on~\citet{griebel1997numerical}) in which two arrays are read with different
data access patterns, across two dimensions.
%%
\begin{minted}[firstnumber=20,xleftmargin=2em]{fortran}
du2dx = ((u(i,j)+u(i+1,j))*(u(i,j)+u(i+1,j))+   &
  gamma*abs(u(i,j)+u(i+1,j))*(u(i,j)-u(i+1,j))- &
  (u(i-1,j)+u(i,j))*(u(i-1,j)+u(i,j))-          &
  gamma*abs(u(i-1,j)+u(i,j))*(u(i-1,j)-u(i,j))) &
  /(4.0*delx)

duvdy = ((v(i,j)+v(i+1,j))*(u(i,j)+u(i,j+1))+   &
  gamma*abs(v(i,j)+v(i+1,j))*(u(i,j)-u(i,j+1))- &
  (v(i,j-1)+v(i+1,j-1))*(u(i,j-1)+u(i,j))-      &
  gamma*abs(v(i,j-1)+v(i+1,j-1))*(u(i,j-1)-     &
  u(i,j))) / (4.0*dely)

laplu = (u(i+1,j)-2.0*u(i,j)+u(i-1,j))/delx/delx+ &
  (u(i,j+1)-2.0*u(i,j)+u(i,j-1))/dely/dely

f(i,j) = u(i,j)+del_t*(laplu/Re-du2dx-duvdy)
\end{minted}
%%
The access pattern is much harder to understand than the one-dimensional
Laplace. The miasma of indexing expressions of the form \texttt{var(i$\pm$a,
j$\pm$b)} is not only hard to read, but is prone to simple textual input
mistakes, \eg{}, swapping \texttt{-} and \texttt{+}, missing an indexing term,
or transforming the wrong variable \eg{} \texttt{(i+1,j)} instead of
\texttt{(i,j+1)}.

In practice, the development process for complex stencil computations involves
ad hoc testing to ensure that no transcription mistakes have been made (\eg{},
by visual inspections, or comparison against an manufactured or analytical
solutions~\cite{farrell2010automated}). Typically such testing is discarded
once the code is seen to be correct.

This is not the only information that is often discarded. The ``shape'' of the
indexing pattern is usually the result of choices made in the
numerical-analysis procedure to convert some continuous equations into a
discrete approximation. Rarely are these decisions captured in the source code,
yet the shape of access is usually uniform and is has a perspicuous and concise
description in numerical analysis literature \eg{}, ``\emph{centered in space,
to a depth of 1}'' referring to indexing terms \texttt{a(i)}, \texttt{a(i-1)}
and \texttt{a(i+1)}.

To support the development of correct stencil computations, we propose a
simple, abstract specification language for the data access pattern of
stencils. The shape of the Laplace example stencil is specified in our language
as:
%
\begin{minted}[linenos=false]{fortran}
!=  stencil centered(depth=1, dim=1) :: a
\end{minted}
%
That is, \texttt{a} is accessed with a symmetrical pattern (``centered'') to a
depth of one in each direction in its first dimension. The Navier-Stokes
example has a shape specified as:
%
\begin{minted}[linenos=false,xleftmargin=0em,xrightmargin=0.8em,breakindent=1.9em]{fortran}
!= stencil centered(depth=1,dim=1)*pointed(dim=2)     + centered(depth=1,dim=2)*pointed(dim=1) :: u

!= stencil forward(depth=1,dim=1)                          * backward(depth=1,dim=2) :: v
\end{minted}
%
The specification requires that, over the whole fragment, \texttt{u} is
accessed with a centered pattern to depth of 1 in both dimensions (this is
known as the \emph{five-point stencil}) and \texttt{v} is accessed in a
neighbourhood bounded forwards to depth of $1$ in the first dimension and
backward to a depth of $1$ in the second dimension. The specification is
relatively small and much more abstract compared to the source code.

We provide a verification tool which can check the spatial correctness of
stencil computations against a specification. Furthermore, the tool can
automatically infer specifications and insert these as comments into the source
code. The main features and benefits of our approach are as follows:
%
\begin{itemize}
\item The primary use case is that a programmer writes a specification first
  before the stencil code. Our tool then checks conformance between the two.
  This specify-and-check approach reduces testing effort and future-proofs
  against bugs introduced during refactoring and maintenance.
%
\item A specification concisely captures the access pattern of a stencil and
  turns it into documentation. The inference procedure efficiently produces
  specifications with no programmer intervention, automatically inserting
  specifications at appropriate places in the source code.
%
\item Our specification format is deliberately abstract, with a small number of
  combinators that \emph{do not involve any indexing expressions}, \eg{}
  \texttt{a(i+1,j-1)}. This contrasts with other specification approaches,
  \eg{}, deductive verification tools such as ACSL, where specifications must
  also be in terms of array-indexing expressions. Therefore, any low-level
  mistakes that could be made whilst programming complex indexing code could
  also be made when writing its specification. Our specifications are more
  abstract and lightweight with the aim of aiding adoption by scientists and
  preventing indexing errors.
\end{itemize}

Our verification tool does not target a class of bugs that can be detected
automatically (\emph{push-button verification}). Instead, stencil computation
bugs must be identified relative to a specification of the intended access
pattern.

In this paper, we solely focus on the definition and the usage of the language
and the acoompanying verification tool and omit the underlying mathematical
model and implementation details due to space considerations. We make the
following contributions:
%
\begin{itemize}
\item We introduce a specification language for stencils that captures common
  forms of data access patterns (\cref{sec:lang});
%
\item We provide an implementation of our approach as an extension to
  CamFort~\cite{camfort}, an open-source program analysis tool for Fortran;
%
\item We report on a quantitative study of stencil computations on a corpus of
  numerical Fortran program (\cref{sec:evaluation}), totalling one million
  lines of code. Our tool identifies and infers specifications for roughly
  60,000 stencil computations in the corpus. Approximately 5,000 of the
  stencils we found are non-trivial, corresponding to code which is a possible
  source of errors.
%
\item We give a verification case study of a stencil computation commonly used
  in scientific computing, simulating programming errors (\Cref{sec:jacobi}).
  Our approach detects all possible simulated programming errors for this
  problem.
\end{itemize}
%

\newcommand{\domainVal}{\mathbb{Z}_\infty}
\section{Stencil specifications}\label{sec:lang}

Our specification language is based on the hypothesis that most forms of array
access in numerical code have a fixed, statically-determined access pattern.
For example, the \emph{five-point stencil} on a two-dimensional array reads
from array indices $(i, j)$, $(i-1, j)$, $(i+1, j)$, $(i, j-1)$, and $(i, j+1)$
for all $i, j$ within the inner boundary of the array (to avoid out-of-bounds
access at the edges). We revisit this hypothesis in \cref{sec:evaluation} with
the inference of such regular stencil patterns on our corpus of numerical
programs.

We give various definitions used in the paper then \cref{subsec:syntax}
outlines the syntax of stencil specifications. \cref{sec:eqs} then provides a
syntactic equational theory for specifications which is sound with respect to
the underlying model.

\paragraph{Notation}
%\label{sec:notation}

\renewcommand*{\arraystretch}{0.8}
For the target language, $e$ ranges over expressions (may be impure) and $v$
over its (imperative) variables. %By
%``variable'', we mean in the imperative sense (named binders to memory cells).

\begin{definition}[Induction variable]
  An integer variable is an \emph{induction variable} if it is the control
  variable of a ``\texttt{for}'' loop incremented by $1$ per iteration. A variable
  is interpreted as an induction variable only within the scope of the loop
  body. Throughout, $i, j, k$ range over induction variables.
%TODO: consider mentioning derived, non-based induction variables
% as we did before.
\end{definition}

\begin{definition}[Array subscripts and indices]
  An \emph{array subscript}, denoted $a(\bar{e})$, is an expression
  which reads from an array $a$ at an \emph{index} specified by a
  comma-separated sequence of integer expressions denoted $\bar{e}$ or
  in expanded form as $(e_1, \ldots, e_n)$. An index $e_i$
  is called \emph{relative} if the expression involves an induction
  variable.
  An \emph{absolute index} is a constant integer expression relative to the
  enclosing loop.
\end{definition}

\begin{definition}[Neighbourhood index]
\label{def:neighbour}
  For an array subscript $a(\bar{e})$ an index $e \in \bar{e}$ is a
  \emph{neighbourhood index} if $e$ is of the form $e \equiv i$, $e \equiv i +
  c$, or $e \equiv i - c$, where $c$ is an integer constant. That is, a
  neighbourhood index is a constant translation of an induction variable. (The
  relation $\equiv$ identifies terms up-to commutativity of $+$ and the inverse
  relationship of $+$ and $-$ \eg{}, $(-b) + i \equiv i - b$).
\end{definition}

\texttt{BEGIN: NEEDS REWRITING}
\begin{definition}[Stencil computations]
\label{def:stencil}
  Let $a$ be an array of dimensionality $n$ and $\overline{b}$ a collection of
  arrays of arbitrary (possibly differing) dimensionalities.  A \emph{stencil
  computation} comprises an iteration by a set $I$ of at most $n$ induction
  variables over a subset of the index space of $a$. Each iteration determines
  the elements of $a$ by an assignment $a(\bar{e}) = e_r$ where each $e \in
  \bar{e}$ is either a neighbourhood or constant index.  For all $b \in
  \bar{b}$ and array subscripts $b(\bar{e}')$ that flow to $e_r$, each $e' \in
  \bar{e}'$ is either a neighbourhood or absolute index.

  We refer to the assignment $a(\bar{e}) = e_r$ and all the associated
  statements with dataflow to $e_r$ as the \emph{stencil kernel}.
\end{definition}

We informally describe what it means for a stencil kernel to be consistent with
a specification. This will be algorithmically described in \Cref{sec:analysis}.
Specifications are associated to a stencil kernel and are declared for one or
more array variables $\bar{v}$ within that kernel.  A stencil specification
defines a \emph{region} associated to an array variable $a$ against which a
stencil kernel should be \emph{consistent}. The region defined by a
specification is an $n$-dimensional rectilinear shape in a discrete space
crossing the origin (which represents a collection of indices which are just
induction variables). Consistency is then a two way requirement. All array
subscripts on $a$ contributing to the kernel should be consistent with the
region. Conversely, all parts of the region should have an index corresponding
it.

\texttt{END: NEEDS REWRITING}

\begin{figure}[t]
\vspace{-0.9em}
\begin{align*}
\def\arraystretch{1}
\setlength{\arraycolsep}{0.2em}
\newcommand{\dimTy}{\mathbb{N}_{>0}}
\begin{array}{rl}
\nonterm{specification} ::= & \nonterm{regionDec} \mid \nonterm{specDec} \\
\nonterm{specDec} ::= & \term{stencil} \; \nonterm{spec} \;
                        \texttt{::} \; v \\
\nonterm{regionDec} ::= &  \texttt{region} \; \texttt{::} \; \nonterm{rvar} \; \texttt{=} \;
                         \nonterm{region}\\[0.4em]
%\nonterm{spec} ::= & \nonterm{spatial} \mid \nonterm{temporal}
%\\[1em]
\nonterm{spec} ::= & [\nonterm{mult},] \; [\nonterm{approx},] \; \nonterm{region} \\
\nonterm{mult} ::= & \term{readOnce} \\
\nonterm{approx} ::= & \term{atMost} \; \mid \; \term{atLeast} \\[0.1em]
\nonterm{region} ::= & \nonterm{rvar} \; \\
\multicolumn{2}{l}{\qquad \mid \; \stenRefl{\dimTy}} \\
\multicolumn{2}{l}{\qquad \mid \;\stenFwd{\mathbb{N}_{>0}}{\dimTy}{\;[, \texttt{nonpointed}]}} \\
\multicolumn{2}{l}{\qquad \mid \; \stenBwd{\mathbb{N}_{>0}}{\dimTy}{\;[, \texttt{nonpointed}]}} \\
\multicolumn{2}{l}{\qquad \mid \; \stenCen{\mathbb{N}_{>0}}{\dimTy}{\;[, \texttt{nonpointed}]}} \\
\multicolumn{2}{l}{\qquad \mid \; \nonterm{region} \, \term{+}
  \, \nonterm{region} \; \mid \; \nonterm{region} \; \term{*} \; \nonterm{region}} \\[0.25em]
%\multicolumn{2}{l}{\qquad\qquad \mid \; \nonterm{rvar}}
%\\[0.5em]
%\nonterm{temporal} ::= \; & \term{dependency} \; (v \; \{ , v \}) [, \texttt{mutual}]
%  \\[0.5em]
\nonterm{rvar} ::= \; & [\text{\term{a}-\term{z}$\,$\term{A}-\term{Z}$\,$\term{0}-\term{9}}]+\\[-1em]
\end{array}
\end{align*}
\caption{Specification syntax (EBNF grammar)}
\label{fig:syntax}
\vspace{-0.8em}
\end{figure}

\subsection{Specification syntax}
\label{subsec:syntax}

\Cref{fig:syntax} gives the syntax of stencil specifications, which is detailed
below. The entry point is the \nonterm{specification} production which splits
into either a \emph{region declaration} or a \emph{specification declaration}.
Regions comprise \emph{region constants} which are combined via region
operators \term{+} and \term{*}.

%\paragraph{Region constants}

Region constants specify a finite contiguous region in a single dimension
relative to the origin and are either \term{pointed}, \term{forward},
\term{backward}, or \term{centered}. The region names are inspired by the
scientific terminology, \eg{} the standard explicit method for approximating
PDEs is known as the \emph{Forward Time, Centered Space} (FTCS)
scheme~\citep{dawson1991finite}.

Each region constant has a dimension identifier $d$ given as a positive natural
number.  Each constant except \term{pointed} has a depth parameter $n$ given as
a positive natural number; \term{pointed} regions implicitly have a depth of
$0$.

A \term{forward} region of depth $n$ specifies a contiguous region in dimension
$d$ starting at the origin. This corresponds to specifying neighbourhood
indices in dimension $d$ ranging from $i$ to $i + n$ for some induction
variable $i$. Similarly, a \term{backward} region of depth $n$ corresponds to
contiguous indices from $i$ to $i - n$ and \term{centered} of depth $n$ from $i
- n$ to $i + n$. A \term{pointed} stencil specifies a neighbour index $i$. For
example, the following shows four specifications with four consistent stencil
kernels reading from arrays \term{a}, \term{b}, \term{c} and \term{d}:
%%
\begin{minted}{fortran}
!= stencil backward(depth=2,dim=1) :: a
e(i, 0) = a(i, 0) + a(i-1, 0) + a(i-2, 0)
!= stencil forward(depth=2,dim=1)*pointed(dim=2) :: b
e(i, j) = b(i, j) + b(i+1, j) + b(i+2, j)
!= stencil centered(depth=1,dim=2) :: c
e(i, j) = (c(j-1) + c(j) + c(j+1))/3.0
!= stencil pointed(dim=3) :: d
e(i, j) = d(0, 0, i)
\end{minted}
%%
Every dimension with neighbourhood indices need to be specified. The remaining
non-conforming indices are not included in the specification, \eg{},
specifications on lines $1$ and $7$ leave some dimensions unspecified since they
do not conform to definition neighbourhood indices.

The \term{forward}, \term{backward}, and \term{centered} regions may all have
an additional attribute \term{nonpointed} which marks absence of the origin.
For example, the following is a \term{nonpointed} \term{backward} stencil
%
\begin{minted}{fortran}
!= stencil backward(depth=2, dim=1, nonpointed) :: a
b(i) = a(i-1) + 10*a(i-2)
\end{minted}

\paragraph{Combining regions}

% r+s means that indices must be consistent with either r or s
% r*s means that indices must be consistent with both r and s at the
% same time

% specifications correspond to a set of schemes

% r+s means an index must be generalised by a scheme in either r or s
%       and every scheme in r must have an index which specialises it
%       and every scheme in s must have an index which specialises it

% r*s means an index must be generalised by a scheme in r and a scheme in s
%       and every pairs of schemes in r and s must have an index which
%           specialises both
%

The region operators \term{+} and \term{*} respectively combine regions
permissively and restrictively.

The restrictive combination of two regions, $r \term{*} s$, means that any
indices in the specified code must be consistent with both $r$ and $s$
simultaneously. Dually, for every pair of components in the regions $r$ and $s$
there must be an index consistent with both. For example, the following
\emph{nine-point stencil} has a specification given by the product of two
\texttt{centered} regions in each dimension:
%%
\begin{minted}[breakindent=2.9em]{fortran}
x = a(i, j)   + a(i-1, j)   + a(i+1, j)
y = a(i, j-1) + a(i-1, j-1) + a(i+1, j-1)
z = a(i, j+1) + a(i-1, j+1) + a(i+1, j+1)
!= stencil centered(depth=1, dim=1) * centered(depth=1, dim=2) :: a
b(i, j) = (x + y + z) / 9.0
\end{minted}
%%
This pattern is common in image convolution applications. The specification
ranges over the values that flow to the array subscript on the left-hand side,
and so ranges over the intermediate assignments to \term{x}, \term{y}, and
\term{z}. Each index in the code is consistent with both specifications
simultaneously, \eg{}, \texttt{a(i-1, j+1)} is within the centered region in
dimension $1$ and the centered region in dimension $2$.

The product $r \term{*} s$ can also be thought of as a bounding box over the
two regions $r$ and $s$.

The permissive combination of two regions $r \term{+} s$ means that any indices
in the specified code must be consistent with either of $r$ or $s$. Dually,
every part of region $r$ must have a corresponding index and similarly for $s$
independently. For example, the following gives the specification of a
five-point stencil which is the sum of two compound \texttt{pointed} and
\texttt{centered} regions in each dimension:
%
\begin{minted}{fortran}
!= stencil centered(depth=1, dim=1)*pointed(dim=2)    + centered(depth=1, dim=2)*pointed(dim=1) :: a
b(i,j) = -4*a(i,j) + a(i-1,j) + a(i+1,j) &
                   + a(i,j-1) + a(i,j+1)
\end{minted}
%%
Here the left-hand side of \texttt{+} says that when the second dimension
(induction variable $j$) is fixed at the origin, the first dimension (induction
variable $i$) accesses the immediate vicinity of the origin (to depth of one).
The right hand side of \texttt{+} is similar but the dimensions are reversed.
This reflects the symmetry under rotation of the five-point stencil. As
presented in \cref{sec:evaluation}, this symmetry is common to stencil
computations used in practice, which makes it easier to write specifications.

\paragraph{Region declarations and variables}

Region specifications can be assigned to region variables via
region declarations. For example, the shape of a
``\emph{Roberts cross}'' edge-detection convolution~\cite{davis1975survey}
can be stated:
%%
\begin{minted}{fortran}
!= region :: r1 = forward(depth=1, dim=1)
!= region :: r2 = forward(depth=1, dim=2)
!= region :: robertsCross = r1*r2
!= stencil robertsCross :: a
\end{minted}
This is useful for common stencil patterns, such as the five-point
pattern, as the region can be defined once and reused.
%%
\paragraph{Modifiers}
%%
Region specifications can be modified
by \emph{approximation} and \emph{multiplicity} information
(in \textit{spec} in \Cref{fig:syntax}).

The \texttt{readOnce} modifier enforces that no index appears more
than once (that is, its multiplicity is one). For example, all of
the previous examples could have \texttt{readOnce} added:
%
\begin{minted}{fortran}
!= stencil readOnce, backward(depth=2, dim=1) :: a
b(i+1) = a(i) + a(i-1) + a(i-2)
\end{minted}
%
This specification would be invalid if any of the array subscripts were
repeated. This modifier provides a way to rule out any accidental repetition of
array subscripts.  The notion is similar to that of linear
types~\cite{wadler1990linear}, where a value must be used exactly once. We opt
for the more informative and easily understood name \texttt{readOnce}. This
modifier is optional, so it need not be present even if the stencil is linear.

In some cases, it is useful to give a lower and/or upper bound for a stencil.
This can be done using either the \term{atMost} or \term{atLeast} modifiers.
This is particularly useful in situations where there is a non-contiguous
stencil pattern, which cannot be expressed precisely in our syntax. For example:
%
\begin{minted}{fortran}
!= stencil atLeast, pointed(dim=1)         :: a
!= stencil atMost, forward(depth=4, dim=1) :: a
b(i) = a(i) + a(i+4)
\end{minted}

\section{Equational theory}
\label{sec:eqs}

Specifications have a theory of equality, $\equiv$, providing axiomatic
semantics for the language. This enables users to rewrite their specification
and provides a way to automatically simplify specifications after inference.
Here we only focus on the former and omit mathematical implications of theory.
Additionally, specifications have a theory of approximation, $<:$, that can be
used by the users to weaken specifications for succinctness in the source.

The equality relation $\equiv$ is defined over \emph{region} syntax which gives
the following axioms:
%
\begin{itemize}
\item $\term{+}$ is idempotent, commutative, and associative;
\item $\term{*}$ is commutative, associative, and distributes over
$\term{+}$, \ie{},  $R \; \texttt{*} \; (S \; \texttt{+} \; T) \; \equiv \; (R \;
                       \texttt{*} \; S) \; \texttt{+} \; (R
                       \; \texttt{*} \; T)$.

\item (\emph{overlapping}) Given two region constants $R_1$ and $R_2$
on the same dimension, then some combinations of region constants
can be coalesced into a single region constant $R_3$ if $R_1$ completely
overlaps $R_2$. There are five possible instantiations of the
following rule $R_1 + R_2 \equiv R_3$:
%
\newcommand{\stenGR}[4]{\small{\textit{#4}(\term{depth=}#1,
  \term{dim=}#2, #3)}}
\newcommand{\stenG}[3]{\small{\textit{#3}(\term{depth=}#1,
  \term{dim=}#2)}}
%
\vspace{-0.2em}
\begin{align*}
\hspace{-1em} & \stenGR{n}{d}{p \vee q}{$R_3$} \\[-0.4em]
\hspace{-1em} \equiv \; & \stenGR{n}{d}{p}{$R_1$} \; \texttt{+} \; \stenGR{m}{d}{q}{$R_2$}
\end{align*}
where $n \geq m$ and $p$ and $q$ range
over booleans representing the absence (true) or presence (false)
of \irrefl{} which are combined via disjunction.
We denote the possibilities as $\term{fwd+fwd} \equiv \term{fwd}$, $\term{bwd+bwd} \equiv \term{bwd}$,
$\term{cen+cen} \equiv \term{cen}$, $\term{cen+fwd} \equiv \term{cen}$ and
$\term{cen+bwd} \equiv \term{cen}$. %Commutative of $\term{+}$
%provides symmetric variants.

\item (\textit{overlapping pointed}) \term{pointed} regions coalesce
  with other regions of the same dimension:
%
\vspace{-0.4em}
\begin{align*}
& \stenG{n}{d}{region} \\[-0.4em]
\equiv \; & \stenGR{n}{d}{p}{region} \; \texttt{+} \;
                          \stenReflS{d}
\end{align*}
\vspace{-1.5em}

\item \textit{(centered)} a forward and backward region of the same
  depth and dimension coalesces into a centered region:
%
\vspace{-0.4em}
\begin{align*}
\hspace{-1em} & {\small{\stenCen{n}{d}{, p \vee q}}} \\[-0.4em]
\hspace{-1em} \equiv \; & {\small{\stenFwdSR{n}{d}{p} \; \texttt{+} \;
                          \stenBwdSR{n}{d}{q}}}
\end{align*}
\vspace{-1.5em}

\item $(\textsc{point}) \;\; (R \, \term{*} \, S^\irreflS) \, \term{+} \,
                         (R^\irreflS \, \term{*} \, S) \; \equiv \; R
                         \, \term{*} \, S$
meaning that for the sum of products of two regions $R$ and $S$, where in each
component one of $R$ or $S$ has the non-pointed attribute denoted by
$R^{\term{np}}$ and $S^{\term{np}}$, the non-pointed attributes cancel to give
$R \, \term{*} \, S$.
\end{itemize}
%
Stencil specifications in \textit{spec} are considered equal when they have the
same modifiers and $\equiv$ equates their regions.

\Cref{fig:inequations} defines a notion of sub-specifications via the relation
$<:^r$ on \textit{region} syntax and then $<:$ on \textit{spec} syntax
inductively.  Both relations are reflexive and transitive (we omit these
equations). The $<:^r$ relation is congruent with respect to \term{+} and
\term{*} and $<:$ is congruent with respect to \term{readOnce}, the equations of
which we omit for brevity.

The \trule{\textsc{eq}} rule connects region equations to approximation;
\trule{\textsc{over}} explains the notion of spatial over-approximation via
overlapping regions; \trule{\textsc{rep}} allows the \term{readOnly} modifier
that to be dropped. The \trule{\textsc{shrink}} and \trule{\textsc{grow}} rules
turn the spatial approximation on regions into approximation on specifications
via the \term{atLeast} and \term{atMost} modifiers. Finally,
\trule{\textsc{Gen$\ast$}} shows that specifications can be generalised by
removing parts of a product $\ast$ which have disjoint dimensions to the rest of
the region product (where $\mathsf{dims}$ denotes the set of dimensions
specified in a region).

\begin{figure}[t]
\vspace{-0.5em}
\begin{align*}
\hspace{-0.5em}
{\small{
\begin{array}{c}
%%
%% INEQUATIONS ON REGION
%%
%\framebox{$\textit{region} <: \textit{region}'$} \\[0.85em]
%
%\dfrac{}{R <:^r R}\trule{\textsc{refl}} \qquad \dfrac{P <:^r R \quad R <:^r Q}{P <:^r
%  Q}\trule{\textsc{trans}}
%\qquad
%\\[1.5em]
\vspace{-0.4em}
\setlength{\arraycolsep}{0.1em}
\dfrac{\hspace{3em} m \leq n \qquad\qquad r \implies s \hspace{3em}}
{\begin{array}{rl}
\stenFwdSR{m}{d}{r} & <:^r \stenFwdSR{n}{d}{s}
%& \;\, ds \subseteq es \wedge n \leq m
\\
\wedge \; \stenBwdSR{m}{d}{r} & <:^r \stenBwdSR{n}{d}{s}
%& \;\, ds \subseteq es \wedge n \leq m
\\
\wedge \; \stenCenSR{m}{d}{r} & <:^r \stenCenSR{n}{d}{s}
%& \;\, ds \subseteq es \wedge n \leq m
\\
\wedge \; \stenBwdSR{m}{d}{r} & <:^r \stenCenSR{n}{d}{s} \\
\wedge \; \stenFwdSR{m}{d}{r} & <:^r \stenCenSR{n}{d}{s} \\[1.2em]
\end{array}}\trule{\textsc{over}} \\[-0.2em]
%\hspace{-0.5em}
%\dfrac{R_1 <:^r S_1 \quad R_2 <:^r S_2}
%      {R_1 \, \texttt{+} \, R_2 <:^r S_1 \, \texttt{+} \, S_2}
%\trule{\textsc{cong}\texttt{+}}
%\dfrac{R_1 <:^r S_1 \quad R_2 <:^r S_2}
%      {R_1 \, \texttt{*} \, R_2 <:^r S_1 \, \texttt{*} \, S_2}
%\trule{\textsc{cong}\texttt{*}} \\[1.5em]
%%
%% INEQUATIONS ON SPEC
%%
%\framebox{$\textit{spec} <: \textit{spec}'$} \\[0.8em]
%
%
\dfrac{R <:^r S}{\term{atLeast}  S <: \term{atLeast}
  R}\trule{\texttt{L}}
\;\;
\dfrac{R <:^r S}{\term{atMost}  R <: \term{atMost}
  S}\trule{\texttt{M}} \\[1.5em]
\dfrac{R <:^r S}{R <: \term{atMost} \, S}\trule{\textsc{grow}} \,
\dfrac{R <:^r S}{S <: \term{atLeast} \, R}\trule{\textsc{shrink}} \,
\dfrac{R \equiv S}{R <:^r S}\trule{\textsc{eq}}
 \\[1.5em]
%%
\dfrac{\mathsf{dims}(R) \cap \mathsf{dims}(S) = \emptyset}{R \ast S <: R}
\trule{\textsc{Gen$\ast$}}
\quad
%%
%\\[1.5em]
%\dfrac{R <: S}{\term{readOnce} \, R <: \term{readOnce} \,
%  S}\trule{\textsc{cong}\texttt{R}}
%\;\;
\;
\dfrac{}{\term{readOnce} \, R <: R}\trule{\textsc{rep}}
%\dfrac{}{R <: \term{atMost} R}
%\quad
%\dfrac{}{R <: \term{atLeast} R}\\[1.5em]
\end{array}}}
\end{align*}
\vspace{-0.5em}
\caption{Approximation $<:^r$ on \textit{regions} and
$<:$ on \textit{specs} (omitting reflexivity, transitivity, and congruences). }
\label{fig:inequations}
\vspace{-0.5em}
\end{figure}

\section{Evaluation}
\label{sec:evaluation}

To study the effectiveness of our approach,
we built a corpus of around 1 million lines of Fortran code from a
range of scientific computing packages: The Unified Model (UM)~\cite{um},
E3MG~\cite{RePEc:aen:journl:2006se-a12}, BLAS~\cite{blas},
Hybrid4~\cite{GBC:GBC635}, GEOS-Chem~\cite{geos-chem}, Navier (based
on \cite{griebel1997numerical}), Computational Physics
2~\cite{nicholas2006computational},
ARPACK-NG~\cite{arpackng}, and
SPECFEM3D~\cite{specfem3d}.

We first examined how frequently stencil computations occur. We parsed
959,427 lines of Fortran code and found that 10\% (97,439) of
statements have a left-hand side as an array subscript on
neighbourhood indices. This supports the idea that stencil-like
computations are common in scientific code.
We then used the inference procedure of the previous section to
generate specifications for stencils in the corpus to assess
the design of the language. %  This assumes that the programs have been implemented
%correctly---
%Normally one would use our approach in verification mode to
%check the correctness of new code.

We would not expect to infer a stencil for each array statement we
found because our analysis restricts
the array-subscript-statements that we classify as a
stencil. In fact, we were able to infer a stencil from 30\% of
these statements. A single statement can involve multiple arrays and
we ended up with 60,525 specifications. This shows that we can express
a large number of stencil shapes within our high-level abstraction and
validates our initial hypothesis that many stencil computations have a
regular shape.

%    numStencilSpecs: 60525
% 4338 files, linesTotal: 1357377, linesParsed: 959427
% Number of stencils (tickAssign) / number of loops
% Percentage of a code base that is taken up by stencils
%Out of 1.35 million lines of code that we fed into our program, and
%the 959,427 lines that we were able to parse, we identified 60,525
%cases that fit our criteria for stencil specification.

%\paragraph{Hypothesis 2: Most stencil computations
%have a regular pattern which can be captured with a simple
%language of contiguous regions}

% Number of stencils we give a spec to (where a spec declaration to n
% variables is counted n times) as percentage of number of stencils

    % tickAssign: 97439
    % tickAssignSuccess: 29402 % <- number of times tickAssign is accompanied by a successful stencil inference
                               % counts only by +1 each time that tickAssign finds a stencil

%In the analysed body of code we found that we were able to deduce
%some form of stencil specification from approximately 30\% of the
%potential stencils examined.

%
%\subsection{Analysing the design of our specifications}

% 1. How may are just reflexive and how many some reflexive
% justReflexive: 55504
% someReflexive: 15823

The majority of specifications generated were relatively simple but we
found significant numbers of more complex shapes. We grouped common
patterns into categories:

\textbf{All pointed} 55,504 of the stencils we found involved
only pointed regions. 39,681 of these were pointed in all
dimensions. Common examples of this were pointwise transformations on
data (such as scaling).

\textbf{Single-action} specifications comprise one
forward, backward, or centered region constant combined via \term{+}
or \term{*} with any number of pointed regions. We identified 4,837 single-action
specifications, of which 1,532 were single-action with a
\texttt{nonpointed} modifier.

\textbf{Multi-action} specifications comprise at least two
 forward, backward, or centered regions, combined with
any number of pointed regions. We identified 265 multi-action
specifications out of which 207 had regions combined only with
$\term{*}$ and 58 combined with a mix of $\term{*}$ and \term{+}.

% multiActionRegionOps1: 31
% multiActionRegionOps2: 115
% multiActionRegionOps8: 1
% multiActionRegionOps14: 5
% multiActionRegionOps17: 4

% %%%%
% Complexity-wise, out of these 156 multi-action specifications, 31 of
% them used a single region operator, 115 used two region operators, 1
% used eight region operators, 5 used fourteen region operators, and 4
% used seventeen region operators. Some of these very complex stencils
% were found in the atmospheric models of the UM.
% %%%%

\textbf{Bounded} specifications occured with 157
\texttt{atMost} bounds and 36 \texttt{atLeast}, the latter of which were
always also paired with an upper bound.

\begin{figure}[t]\begin{minted}[fontsize=\scriptsize,breakindent=0em,linenos=false,xleftmargin=0em,breakafter=)]{fortran}
!=stencil readOnce,(forward(depth=1,dim=3,nonpointed))*(backward(depth=1,dim=1))*(backward(depth=1,dim=2))+(forward(depth=1,dim=3))*(backward(depth=1,dim=1))*(pointed(dim=2))+(forward(depth=1,dim=3))*(backward(depth=1,dim=1,nonpointed))*(backward(depth=1,dim=2,nonpointed))+(forward(depth=1,dim=3))*(backward(depth=1,dim=2))*(pointed(dim=1))+(backward(depth=1,dim=1))*(backward(depth=1,dim=2,nonpointed))*(pointed(dim=3))+(backward(depth=1,dim=1,nonpointed))*(backward(depth=1,dim=2))*(pointed(dim=3))::x
\end{minted}
\caption{Complex specification inferred from
  \textbf{UM}\label{fig:smagorinsky}}
\vspace{-1em}
\end{figure}

The single- and multi-action classes represent more complex stencils
with a real possibility for programmer error. As an extreme example, the Unified
Model has an implementation of the Smagorinsky subgrid-scale model for
calculating turbulence on which our inference yields 39 specs from 340
lines of code. This is a large reduction given
the complexity of the algorithm.  We show one example
in~\Cref{fig:smagorinsky} which specifies the access pattern to
a 3-dimensional array (which we have renamed to
\texttt{x}) arising from a kernel of 93 lines of code involving 142
array subscripts. The specification was the most complex
seen in our corpus, yet it still represents a significant abstraction
of the spatial behaviour given size and complexity of the kernel it describes.

We measured the frequency at which individual specifications involved
multiple occurences of \term{+} and \term{*}:

\vspace{0.25em}
\setlength{\tabcolsep}{0.57em}
{\small{
\hspace{-1em}\begin{tabular}{c|cccccccc}
& 0 & 1 & 2 & 3 & 4 & 5 & 6 & $> 7$ \\ \hline
\term{+} & 60,265 & 220 & 15 & 12 & 5 & 4 & 4 & 0 \\
\term{*} & 31,393 & 15,151 & 13,457 & 424 & 41 & 42 & 5 & 12
\end{tabular}}}
\vspace{0.2em}

\noindent
\eg{}, roughly half the specifications did not involve $\term{*}$,
a quarter use one $\term{*}$ and just under a quarter
use two $\term{*}$ operators.

\paragraph{Limitations}

There were various reasons why we did not infer
specifications on every looped array computation: \\
\noindent
1) \textbf{Non-subset induction variables} occur when the
induction variables on the RHS are not a subset of those in the LHS. These
cases are not stencils by our definition. \\
\noindent
2) \textbf{Derived induction variables} where an
index (\mintinline{fortran}{x}) is derived from an
induction variable (\mintinline{fortran}{i}) as in
\mintinline{fortran}{x = len - i};  \\
\noindent
3) \textbf{Inconsistent induction dimensions} occur when
an induction variable is used to specify more than one array dimension
on the RHS or multiple induction variables are used for the same
dimension on the RHS. These are common in matrix operations such as
LU-decomposition with assignments such as
\mintinline{fortran}{a(l) = a(l) - a(m) * b(l, m)}.

\subsection{Detecting errors in the 2-D Jacobi iteration}
\label{sec:jacobi}
One common example of a stencil computation is the two-dimensional
Jacobi iteration that repeatedly goes through each cell in a matrix
and computes the average value of the four adjacent cells. The kernel
is given by:
\begin{minted}{fortran}
  a(i,j) = (a(i-1,j)+a(i+1,j)+a(i,j+1)+a(i,j-1))/4
\end{minted}
We infer a precise specification of its shape as:
\begin{minted}[breakafter=+:,breakindent=-0.6em,breaksymbolsep=0.4em,linenos=false,xleftmargin=-0.5em]{fortran}
  != stencil pointed(dim=1)*centered(depth=1,dim=2,nonpointed)+pointed(dim=2)*centered(depth=1,dim=1,nonpointed)::a
\end{minted}
%
To test our implementation,
we examined whether programmer errors would be detected by replacing
the array index offsets with $-1$, $0$, or $1$ and running our
verification algorithm. Our checking procedure correctly reported a verification
failure in each of 6,537 permutations corresponding to an error.
The iteration computes the average of four adjacent
cells so $24$ ($4$ factorial) of the possible array index perturbations
are correct, all of which are accepted by our checker.

% /um/trunk/src/atmosphere/diffusion_and_filtering/turb_smagorinsky.F90
% Tried introducing indexing errors replacing one of the indices of cx_rho_instance with -1,0,+1
% 42 cases were caught
% 6 cases were not caught

% (/ 55504.0 60525) = 91.7%
% (/ 15823.0 60525) = 26.1%
% 91.7\% of the specifications involved only reflexive regions. 26.1\%

% 2. Break down of "interesting specifications" (contain more than
% just reflexive)
%   - Single-action, histogram on depths
%   - Single-action + irreflexives, histogram on depths

% singleAction: 4837
% singleActionIrr: 1532

%   - Multi-action combined only by *
%   - Multi-action combined also by +, histogram of tree depths

    % multiAction: 156
    % multiActionMulOnly: 146

%   - Bounded specs (how many lower, upper, or both)

    % atLeast: 36
    % atMost: 157
    % boundedBoth: 36

% Present not just percentages but concrete numbers

\section{Related work and conclusions}
\label{sec:discussion}

Various deductive verification tools can express array indexing
in their specifications, \eg{}, ACSL of \citet{baudin2008acsl}
for C (\eg{}~\citet[Example 3.4.1]{burghardt2010acsl}). A specification
can be given for a stencil computation but must use fine-grained
indexing as in code and therefore is similarly prone to indexing
errors. Our approach is much more abstract-- it does
not aim to reify indexing in the specification, but
provides simple spatial descriptions which capture a large number
of common patterns.

\citet{kamil2016verified} propose \emph{verified lifting} to extract a
functionally-complete mathematical representation of low-level, and
potentially optimised, stencils in Fortran code. This extracted
predicate representation of a stencil is used to generate code
for the \textsc{Halide} high-performance
compiler~\citep{ragan2013halide}. Thus they must capture the
full meaning of a stencil computation which requires significant
program analysis. For example, they report that some degenerate
stencil kernels take up to 17 hours to analyse and others require
programmer intervention for correct invariants to be inferred.

% ACR-I think we could skip this in the interests of space
%This led them to develop \textsf{STNG}, a loop invariant and postcondition
%finder through syntax-guided synthesis. Using this approach they restrict the
%search space of postconditions to predicates of a certain form, one that is
%compatible with \textsf{Halide}. They then look at concrete loop iterations and
%form an hypothesis about the invariant, which is later confirmed or put back
%into the system using a SMT solver.

Our approach differs significantly. Rather than
full representation of stencils, we focus on specifying
just the spatial behaviour in a lightweight way.
Thus, it suffices for us to perform a comparatively
simple data-flow analysis which is efficient, scales linearly with
code size, and does not require any user intervention.
Whilst we do not perform deep semantic analysis of stencils,
 the analysis part of our
approach can be made arbitrarily more sophisticated independent of the rest of
the work.
% Hence, we do not require SMT solving or search of loop invariants. It
%suffices to do comparatively simple data-flow analysis.
%Since CamFort mostly does
%syntactic analysis, the inference procedure terminates quickly and
%never requires programmer intervention.
%The downside of our
%approach is that if optimisation causes the stencil computation to be heavily
%obfuscated, \textsf{STNG} would capture the access behaviour better.
%
Furthermore, Kamil \emph{et al.} do not provide a user-visible
syntactic representation of their specifications, and nor do
they provide verification from specifications \eg{}, to future-proof
the code against later changes. Even if they were to provide a
syntactic representation, for complex stencils such as Navier-Stokes
from \cref{sec:intro}, it would be as verbose as the code itself,
making it difficult for programmer to understand the overall shape of
the indexing behaviour.

% GPGPU
Our work has similarities with efforts to verify kernels
written for General-Purpose GPU
programming, such as in \citet{Blom:2014:SoCP}. %Stencils are a form of
%kernel, and GPGPU programming can be viewed as a massively parallel
%method of transforming a large matrix.
However, their focus is
mainly on the synchronisation of kernels and the avoidance of
data races., while we are interested in correctness
%embedded within a more typical general-purpose programming
%language. %Other work on GPGPU computation, such as
%\citet{Zhang:2012:CGO}, has focused primarily on generating
%optimised code based on relatively simple specifications: to
%provide performance while keeping the programmer's effort within
%reason.
%
% Sketching Stencils
\citet{Solar-Lezama:2007:PLDI} give specifications of stencils using
unoptimised ``reference'' stencils, coupled with partial
implementations which are completed by a code
generation tool. %These kinds of specifications are just simple
%implementations, so this tool is useful for hand-written, optimised
%stencils.

% Pochoir
\citet{Tang:2011:SPAA} define a specification language for writing stencils
embedded in C++ (with Cilk~\citep{blumofe1996cilk} extensions) that are then
compiled into parallel programs based on trapezoidal decompositions with
hyperspace cuts.

By contrast, the work of \citet{Abe:2013:IPDPSW} studies
correctness bringing a form of model-checking to
verify certain stencil computations in the context of
parallelism
 in partitioned global address
space languages. %These are scenarios where an array is divided into
%subarrays on multiple processors but each has the ability to access
%the others' memory. The authors avoid the state explosion dilemma of
%model-checking by relying upon the fact that accesses to the
%``boundary elements'' will require a different method by virtue of the
%non-locality of that memory. Since this access is performed in a
%different manner, those can be identified automatically by static
%analysis. To achieve this,
Abe \emph{et al.} provide a new language for writing stencil computations. Much of the specification
effort goes towards describing the distribution of the
computation over multiple processors. The code for the stencil kernel
is generated from a relatively high-level specification.
In contrast, we integrate directly
into existing, legacy codebases and established languages, bringing
the benefits of verification more easily to scientific computing.%such as
%older versions of Fortran.% We infer stencil specifications from
%Fortran code and check annotations on stencil computations within
%Fortran code.

\paragraph{Concluding remarks}

There is an increasing awareness of the need for verification
techniques in
science~\cite{post2005computational,oberkampf2010verification,orchard2014computational}.
Our specification language is an approach in this direction, providing
a system of lightweight, high-level specification for numerical code.
As opposed to existing approaches our language takes inspiration from
the numerical literature and provides a concise, abstract description
of the stencil.

We evaluated our approach by extending CamFort, an open-source analysis
tool for Fortran programs~\cite{camfort}, running the inference process over a corpus
of a million lines of Fortran code. We showed that our language is
capable of capturing many real-world uses of stencils and showed in a
case-study that it will detect (simulated) programmer errors.

Our restrictive definition of stencils means we do not generate
specifications for a large number of iterated array accesses in the
corpus, \eg{} that incorporate reductions. For
future work, we will look at expanding the flexibility of
our analysis to include these examples.

\bibliography{references}

\end{document}

%%  LocalWords:  refactoring affine parameterised nonpointed atMost
%%  LocalWords:  centered readOnce atLeast discretisation Equational
%%  LocalWords:  equational disjunction denotational dimensionality
%%  LocalWords:  interprocedural Fortran CamFort preprocessor
%%  LocalWords:  committers