### 2017-03-13 oopsla17/paper.tex
oopsla stub

\documentclass[acmlarge,anonymous,review]{acmart}

\title{Understanding and Verifying the Shape of Stencil Computations}

\begin{document}
\maketitle

%\bibliographystyle{

\section{Introduction}

\section{An Empirical Study of Stencils in Scientific Fortran code}

\begin{enumerate}
\item Stencils are ubiquitious;
\item Most stencils read their data from source arrays with a static pattern
base on constant offsets from the mid-point;
\item Most stencils read their data from contiguous \emph{regions} in an array;
\item Most stencils read each index at most once.
\end{enumerate}

\section{A Specification Language for Shape}

\section{Analysis, Checking, and Inference}

\section{Evaluation}

\bibliography{references}

\end{document}