### 2017-04-11 oopsla17/oopsla17.tex
title and abstract tweak

\documentclass[acmlarge,review,anonymous]{acmart}\settopmatter{printfolios=true}

%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption

\usepackage{minted}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{syntax}
\usepackage{indentfirst}
\usepackage{cleveref}
\usepackage{xcolor}
\usepackage{siunitx} % For pretty-printing numeric values and SI units
                     % of measure. e.g., the tabular column type S is
                     % used to print nice-looking tables of numbers.
\sisetup{ % defaults
  group-separator={,},
  group-minimum-digits={3},
  output-decimal-marker={.},
  table-format = 6
}

\usepackage{stencilmacros}

%% Bibliography style
\bibliographystyle{ACM-Reference-Format}
%% Citation style
%% Note: author/year citations are required for papers published as an
%% issue of PACMPL.
\citestyle{acmauthoryear}   %% For author/year citations

\fvset{fontsize=\scriptsize}

\setlength{\grammarindent}{3em} % increase separation between LHS/RHS

\newcommand{\inbar}{\ \ \textbar\ \ }

% For evluation and empirical study
% Defines the commands (headers)
\input{resultsCmds}
% renews the commands with actual data
%\input{results}

\title{Verifying Spatial Properties of Array Computations}

\author{Dominic Orchard}
\affiliation{
  \department{School of Computing}
  \institution{University of Kent}
}
\email{d.a.orchard@kent.ac.uk}

\author{Mistral Contrastin}
\affiliation{
  \department{Computer Laboratory}
  \institution{University of Cambridge}
}
\email{Mistral.Contrastin@cl.cam.ac.uk}

\author{Matthew Danish}
\affiliation{
  \department{Computer Laboratory}
  \institution{University of Cambridge}
}
\email{Matthew.Danish@cl.cam.ac.uk}

\author{Andrew Rice}
\affiliation{
  \department{Computer Laboratory}
  \institution{University of Cambridge}
}
\email{Andrew.Rice@cl.cam.ac.uk}

\begin{abstract}
  Arrays computations are the workhorse of numerical modelling and
  computational science. However, low-level manipulation of array
  indices is a source of program error.  Many practitioners are aware
  of the need to ensure program correctness, yet very few of the
  techniques from the programming research community are applied by
  scientists. We aim to change that by providing targetted lightweight
  verification techniques for scientific code.  We focus on the all
  too common mistake of array offset errors as a generalisation of
  off-by-one errors.  Firstly, we report on a code analysis study on
  twelve real-world computational science code base, identifying
  common idioms of array usage and their spatial properties. This
  provides much needed data on array programming idioms common in
  scientific code.  From this data, we designed a lightweight
  declarative specification language capturing the majority of array
  access patterns via a small set of combinators. We detail a semantic
  model, and the design and implementation of a verification tool for
  our specification language, which both checks and infers
  specifications.  We evaluate our tool on our corpus of scientific
  code and give verification case studies of bug fixes that are
  detected by our approach. We found roughly 60,000 targets for
  specification across roughly 1.2 million lines of code, showing the
  vast majority of array computations read from arrays in a pattern with
  a simple, regular, static shape.
\end{abstract}

%  We focus on one error-prone aspect of numerical code: the 
%  data access pattern of arrays.

\begin{document}
\maketitle

\section{Introduction}

\paragraph{Context of this research}

There is an increasing awareness of the need for program verification
techniques in
science~\cite{post2005computational,oberkampf2010verification,orchard2014computational}.
Whilst there are various kinds of automated and semi-automated
verification tools available, they see little use in science. This is
partly due to a lack of training and a lack of awareness. But its also
due to a lack of targetting. This paper is part of a line of research
to provide lightweight verification tools targetting
programming patterns common in computational science code, 
that are easy to use by scientists, and that integrate with common
practises and workflows.

\subsection{Terminology and notation}

\begin{definition}[Induction variables]
  An integer variable is a \emph{base induction variable} if it is the control
  variable of a ``\texttt{for}'' loop, incremented by $1$ per iteration. A variable
  is interpreted as an induction variable only within the scope of the loop
  body. Throughout, $i, j, k$ range over induction variables.

  A \emph{derived induction variable} is an expression of the form $a
  i + b$, \ie{}, an affine expression on an induction variable $i$.
\end{definition}

\begin{definition}[Array subscripts and indices]
  An \emph{array subscript}, denoted $a(\bar{e})$, is an expression
  which reads from an array $a$ at an \emph{index} specified by a
  comma-separated sequence of integer expressions denoted $\bar{e}$ or
  in expanded form as $(e_1, \ldots, e_n)$. An index $e_i$ is called
  \emph{relative} if the expression involves an induction variable.
  An \emph{absolute index} is a integer expression which is constant
  relative to the enclosing loop.
\end{definition}

\begin{definition}[Neighbourhood index]
\label{def:neighbour}
  For an array subscript $a(\bar{e})$ an index $e \in \bar{e}$ is a
  \emph{neighbourhood index} if $e$ is of the form $e \equiv i$, $e \equiv i +
  c$, or $e \equiv i - c$, where $c$ is an integer constant. That is, a
  neighbourhood index is a constant translation of an induction variable. (The
  relation $\equiv$ identifies terms up-to commutativity of $+$ and the inverse
  relationship of $+$ and $-$ \eg{}, $(-b) + i \equiv i - b$).
\end{definition}

\section{An Empirical Study of Array Computations in Scientific Fortran code}

\begin{enumerate}
\item Stencils are ubiquitious;
\item Most stencils read their data from source arrays with a static pattern
base on constant offsets from the mid-point;
\item Most stencils read their data from contiguous \emph{regions} in an array;
\item Most stencils read each index at most once.
\end{enumerate}

\subsection{Study}
%
\noindent
We put together a corpus of around 1 million lines of Fortran code from a
\numPackages{} scientific computing packages ranging in size and scope: The Unified Model (UM)~\cite{um},
E3MG~\cite{RePEc:aen:journl:2006se-a12}, BLAS~\cite{blas},
Hybrid4~\cite{GBC:GBC635}, GEOS-Chem~\cite{geos-chem}, Navier (based
on \cite{griebel1997numerical}), Computational Physics
2~\cite{nicholas2006computational},
ARPACK-NG~\cite{arpackng}, and
SPECFEM3D~\cite{specfem3d}. Appendix~\ref{app:corpus} provides
further detail on these packages.

\begin{itemize}
% (Vars, ...)
\item An assignment to a variable, with an RHS comprising some array
  computation, tends to correspond to a reduction (e.g., calculating
the max value in an array).

\end{itemize}

\section{A Specification Language for Shape}

\begin{figure}[t]
\begin{align*}
\def\arraystretch{1}
\setlength{\arraycolsep}{0.2em}
\newcommand{\dimTy}{$\mathbb{N}_{>0}$}
\newcommand{\depthRange}{$\mathbb{N}_{>0}$}
\begin{array}{rl}
\nonterm{specification} ::= & \nonterm{regionDec} \mid \nonterm{specDec} \\
\nonterm{specDec} ::= & \term{stencil} \; \nonterm{spec} \;
                        \texttt{::} \; v \\
\nonterm{regionDec} ::= &  \texttt{region} \; \texttt{::} \; \nonterm{rvar} \; \texttt{=} \;
                         \nonterm{region}\\[0.4em]
%\nonterm{spec} ::= & \nonterm{spatial} \mid \nonterm{temporal}
%\\[1em]
\nonterm{spec} ::= & [\nonterm{mult},] \; [\nonterm{approx},] \; \nonterm{region} \\
\nonterm{mult} ::= & \term{readOnce} \\
\nonterm{approx} ::= & \term{atMost} \; \mid \; \term{atLeast} \\[0.1em]
\nonterm{region} ::= & \nonterm{rvar} \; \mid \;
                       \stencil{p}{\dimTy}{}{} \; \\
& {\qquad\; \mid \; \stencil{f}{\dimTy}{\depthRange}{\;[, \nonpointed]}} \\
& {\qquad\; \mid \; \stencil{b}{\dimTy}{\depthRange}{\;[, \nonpointed]}} \\
& {\qquad\; \mid \; \stencil{c}{\dimTy}{\depthRange}{\;[, \nonpointed]}} \\
& {\qquad\; \mid \; \nonterm{region} \, \term{+}
  \, \nonterm{region} \; \mid \; \nonterm{region} \; \term{*} \; \nonterm{region}} \\[0.25em]
%\multicolumn{2}{l}{\qquad\qquad \mid \; \nonterm{rvar}}
%\\[0.5em]
%\nonterm{temporal} ::= \; & \term{dependency} \; (v \; \{ , v \}) [, \texttt{mutual}]
%  \\[0.5em]
\nonterm{rvar} ::= \; & [\text{\term{a}-\term{z}$\,$\term{A}-\term{Z}$\,$\term{0}-\term{9}}]+\\[-1em]
\end{array}
\end{align*}
\caption{Specification syntax (EBNF grammar)}
\label{fig:syntax}
\end{figure}

\subsection{Specification syntax}
\label{subsec:syntax}

\Cref{fig:syntax} gives the syntax of stencil specifications, which is
detailed below. The entry point is the \nonterm{specification} production which
splits into either a \emph{region declaration} or a \emph{specification
declaration}.
Regions comprise \emph{region constants} which are
combined via region operators \term{+} and \term{*}.

\section{Analysis, Checking, and Inference}

\section{Evaluation}

\bibliography{references}

\appendix
% http://2017.splashcon.org/track/splash-2017-OOPSLA#Instructions-for-Authors
% "There is no page limit for bibliographic references and appendices,
% and, therefore, for the overall submission."
\input{appendix}

\end{document}