### 2017-04-17 oopsla17/introduction.tex
fix to coalesce

\paragraph{Context of this research}

There is an increasing awareness of the need for program verification
techniques in
science~\cite{post2005computational,oberkampf2010verification,orchard2014computational}.
Whilst there are various kinds of automated and semi-automated
verification tools available, they see little use in science. This is
partly due to a lack of training and a lack of awareness. But its also
due to a lack of targetting. This paper is part of a line of research
to provide lightweight verification tools targetting
programming patterns common in computational science code, 
that are easy to use by scientists, and that integrate with common
practises and workflows.

\subsection{Terminology and notation}

\begin{defn}[Induction variables]
  An integer variable is a \emph{base induction variable} if it is the control
  variable of a ``\texttt{for}'' loop, incremented by $1$ per iteration. A variable
  is interpreted as an induction variable only within the scope of the loop
  body. Throughout, $i, j, k$ range over induction variables.

  A \emph{derived induction variable} is an expression of the form $a \times i
  + b$, where $a$ and $b$ are constant expressions, \ie{}, an affine expression
  on an induction variable $i$.
\end{defn}

\begin{defn}[Array subscripts and indices]\label{def:array-subs}
  An \emph{array subscript}, denoted $a(\bar{e})$, is an expression which reads
  from an $N$-dimensional array $a$ at an \emph{index} specified by a
  comma-separated sequence of integer expressions denoted $\bar{e}$ or in
  expanded form as $(e_1, \ldots, e_N)$. An index $e_i$ is called
  \emph{relative} if the expression involves an induction variable. An
  \emph{absolute index} is a integer expression which is constant relative to
  the enclosing loop.
\end{defn}

\begin{defn}[Neighbourhood index]\label{def:neighbour-ix}
  For an array subscript $a(\bar{e})$ an index $e \in \bar{e}$ is a
  \emph{neighbourhood index} if $e$ is of the form $e \equiv i$, $e \equiv i +
  c$, or $e \equiv i - c$, where $c$ is an integer constant. That is, a
  neighbourhood index is a constant translation of an induction variable. (The
  relation $\equiv$ identifies terms up-to commutativity of $+$ and the inverse
  relationship of $+$ and $-$ \eg{}, $(-b) + i \equiv i - b$).
\end{defn}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: