### 2017-04-18 oopsla17/model.tex

\noindent
We define a lattice model of array access patterns which serves as
a semantic model of our specification language and a model of array
access patterns in source code. This model (1) used to explain the
meaning of our specifications; (2) as a basis for the inference and
checking algorithms (\Cref{subsec:analysis}); (3) justifies an equational
theory for specifications in the next section; (4) is used to optimise
specifications using lattice identities; and (5) can be used to guide
correct implementations.

The model is defined over vectors of sets of integers which we call
\emph{index schemes}.  As an initial informal example, consider the
a stencil computation with loop body \fortran{y(i) = x(i,0) + x(i+1,0)}.
The access pattern on array \term{x}, relative
to induction variable \term{i}, is captured
by a vector of length 2 containing integer sets $\langle{\{0, 1\},
  \mathbb{Z}\}}\rangle$. This describes that, in the first dimension,
the array is read at offsets of $0$ and $1$ from an induction
variable. In the second dimension, the index is unconstrained
as it is a constant.

Index schemes form a lattice which provides a rich set of equations
and properties, which we exploit. We first set up the 
domain of the model (\S\ref{sec:domain}), using it to define a semantics
for our specification language (\S\ref{sec:semantics}) and then as
the target for an interpretation on imperative code (\S\ref{sec:fromcode}).
We state various results in this section, for which the proofs are
provided in \Cref{sec:proofs}.

\subsection{Lattice model of regions}
\label{sec:domain}

%\begin{defn}[Extended integers]
%  We define the set \zinf{} as
%  $\mathbb{Z}$ extended with $\infty$ and $-\infty$. For any $a$ in
%  \zinf{}, we have $-\infty \leq a \leq \infty$. The resulting set is
%  a total order with top and bottom elements $-\infty$ and
%  $\infty$ respectively.
%\end{defn}

\begin{defn}[Index scheme]
  An $N$-dimensional \emph{index scheme} is a vector of length $N$ of
  integer sets, \ie{}, a member of $\mathcal{P}(\bz)^{N}$.
  An equivalent view of these vectors is as an $N$-times Cartesian
  product on subsets of $\bz$.
  We use $S, T, U$ to denote index schemes. Henceforth, we assume index schemes
  are all $N$-dimensional\footnote{The particular
    dimensionality is derived from array types when the model is
    used.} for some $N$.

  Index schemes can be \emph{projected} in the $i^{\textit{th}}$ dimension by
  $\pi_i : {\mathcal{P}(\bz{})}^N \to \mathcal{P}(\bz{})$. For an index scheme
  $I$, we refer to $\pi_i(S)$ as the \emph{$i^{th}$ component} of $S$. We assume
  that $i$, when used for projection, always lies between $1$ and $N$.
\end{defn}

\begin{restatable}{lem}{vectorIntersect}
\label{lem:vector-intersect}
  Intersection distributes over index schemes. That is, for index schemes $S, T
  \in \mathcal{P}(\bz)^N$
%
  \begin{equation*}
    S \cap T = \prod_{i = 1}^{N} \pi_i(S) \cap \pi_i(T)
  \end{equation*}
\end{restatable}

Union does not distribute over index schemes, however, a more restricted
property holds.

\begin{restatable}{lem}{vectorUnion}
\label{lem:vector-union} If
$S$ and $T$ are index schemes where $\pi_i(S) = \pi_i(T)$ for all $1 \leq i
\leq N$ apart from some dimension $k$, then:
%
  \begin{equation*}
    S \cup  T
    =
    \pi_1(S) \times \cdots \times
    (\pi_k(S) \cup \pi_k(T)) \times \cdots \times
    \pi_N(S)
  \end{equation*}
\end{restatable}
%
\begin{defn}[Intervals with an optional hole]
  We define an extended notion of closed interval on $\bz{}$ which may contain a
  \emph{hole} at the origin, written \interv{a}{b}{c} where $a$ and $b$ are
  drawn from \bz{} with $a \leq 0 \leq b$ and $c$ is drawn from $\mathbb{B} = \{
  \mathsf{true}, \mathsf{false} \}$. Intervals are interpreted as sets as
  follows:
%
  \begin{equation*}
    \interv{a}{b}{c} \triangleq
      \{ n \mid a \leq n \leq b \wedge (\neg c \implies n \neq 0) \}
  \end{equation*}
  %
  If the
  superscript to the interval is omitted it is treated as $\mathsf{true}$.
  We also add the distinguished interval $\interv{-\infty}{\infty}{}$, which is simply an
  alias for \bz{}, but this notation prevents handling infinite interval
  separately in the following definitions, lemmas, and proofs. Here, $-\infty$
  and $\infty$ behave as top and bottom elements to \bz{} respectively.

  We denote the set of all such intervals as $\textit{Interval}$. 
\end{defn}
%
\begin{restatable}{lem}{intervalIdentities}
 \label{lem:interval-identities}
  We have the following dual identities for \bz{} intervals:
%
  \begin{align*}
    \interv{a}{b}{c} \cap \interv{d}{e}{f} & =
      \interv{\max \{a,d\}}{\min \{b,e\}}{c \wedge f} \\
    \interv{a}{b}{c} \cup \interv{d}{e}{f} & =
      \interv{\min \{a,d\}}{\max \{b,e\}}{c \vee f}
  \end{align*}
\end{restatable}
%

We define two specialisations of index schemes:
\emph{subscript scheme} and \emph{interval scheme}.

\begin{defn}[Interval scheme]\label{def:interval-scheme}
  An interval scheme is an $N$-length vector ($N$-ways Cartesian
  product) of intervals on \bz{}, denoted by the set
  $\textit{Interval}^N$ for a product of $N$ intervals.
\end{defn}

\begin{defn}[Subscript scheme]
  A \emph{subscript scheme} is an index scheme scheme where:
  %
  \begin{equation*}
    \forall i.\ 1 \leq i \leq N \implies
      \pi_i (S) = \{ p \}
      \; \vee \;
      \pi_i (S) = \interv{-\infty}{\infty}{}
    \end{equation*}
%
  That is, the $i^{th}$ component of the set is either a singleton in \bz{} or
  the infinite interval.
\end{defn}
%
\begin{defn}[Region]
  A region is an index scheme and \region{N} is the set of all regions (\ie{},
  $\region{N} \subseteq \bz^N)$. The set of all regions is  defined
  as the smallest set satisfying the following:
%
  \begin{enumerate}
    \item If $R$ is in $\textit{Interval}^N$, then $R$ is in \region{N}.
    \item If $R$ and $S$ are in \region{N}, then so are $R \cap S$ and
      $R \cup S$.
  \end{enumerate}
\end{defn}
%
\begin{restatable}{prop}{regionLattice}
  \label{prop:regionLattice}
  $(\region{N},\cup,\cap,\subseteq)$ is a bounded distributive lattice with top
  $\top = \bz{}^N$ and bottom $\bot = \emptyset$.
\end{restatable}
%
The set of regions $\region{N}$ is target of our specification
language model, and of the code analysis. We further wrap this in a
labelled variant which provides information on multiplicity and approximation.

\begin{defn}
  \label{defn:modifiers}
  $\mathsf{Mult}$ and $\mathsf{Approx}$ are parametric labelled variant types
  with injections given by their definition:
%
  \begin{align*}
    \mathsf{Mult} \;\; a \;\,
      \triangleq \;\, \mathsf{mult} \; a \;\mid\; \mathsf{only} \; a \qquad\quad
    \mathsf{Approx} \;\; a \;\,
      \triangleq \;\, \mathsf{exact} \; a \;\mid\; \mathsf{lower} \; a \;\mid\;
        \mathsf{upper} \; a
  \end{align*}
%
  \eg{}, $\mathsf{lower}$ is an injection $\mathsf{lower} : a \to \mathsf{Approx}
  \; a$ etc.
%  These will be used in the following subsection to give meaning to the
%  specification modifiers for approximation and multiplicity.
\end{defn}

\subsection{Denotational semantics for specifications}
\label{sec:semantics}

\noindent
An interpretation function $\interp{-}_N$ maps closed\footnote{That
  is, we assume there are no occurrences of \textit{rvar} in a
  specification being modelled.  Any \emph{open} specification
  containing region variables can be made closed by straightforward
  syntactic substitution with a (closed) \textit{region}.}
specifications to sets of $N$-dimensional index schemes with modifier
information, \ie{} specifications are mapped to
$\textsf{Mult} (\textsf{Approx} (\region{N}))$.

The interpretation is overloaded on \emph{regions} in
\Cref{subfig:region-model}. Various intermediate notions are used.

\begin{defn}
  Let $\textit{promote}_N : \mathbb{N}^+ \times \textit{Interval} \to
  \textit{Interval}^N$ be a function generating an interval scheme such that if
  $v$ is $\vecgen{N}{i}{\interv{a}{b}{c}}$, then $\pi_i(v) = \interv{a}{b}{c}$
  and $\pi_j(v) = \bz{}$ in all other dimensions $j$.
\end{defn}

%
\begin{figure}[!t]
 \label{fig:semantics}
\begin{subfigure}[t]{0.4\textwidth}
\begin{align*}
  \interp{-}_N & : \textit{region} \rightarrow \region{N} \\
  %
   \interp{\texttt{\textcap{r} + \textcap{s}}}_N & =
    \interp{\textcap{r}}_N \vee \interp{\textcap{s}}_N
\\
  \interp{\texttt{\textcap{r} * \textcap{s}}}_N & =
   \interp{\textcap{r}}_N \wedge \interp{\textcap{s}}_N \\
%%
  \interp{\texttt{\textcap{rconst}}}_N & =
    \vecgen{N}{i}{\interp{\texttt{\textcap{rconst}}}^{\mathsf{c}}}
  \\[1em]
  %%
  \interp{-}^{\mathsf{c}} & : \textit{rconst} \rightarrow
                            \textit{Interval} \\
  \interp{\stencil{p}{$i$}{}{}}^{\mathsf{c}} & =
    {\interv{0}{0}{\mathsf{true}}}\\
%
  \interp{\stencil{c}{$i$}{$k$}{\textcap{p}}}^{\mathsf{c}} & =
    {\interv{-k}{k}{\interp{\textcap{p}}}} \\
%
  \interp{\stencil{f}{$i$}{$k$}{\textcap{p}}}^{\mathsf{c}} & =
    {\interv{0}{k}{\interp{\textcap{p}}}} \\
%
  \interp{\stencil{b}{$i$}{$k$}{\textcap{p}}}^{\mathsf{c}} & =
    {\interv{-k}{0}{\interp{\textcap{p}}}}
\end{align*}
\caption{Interpretation of regions}
\label{subfig:region-model}
\end{subfigure}
\hspace{1em}
\begin{subfigure}[t]{0.3\textwidth}
\begin{align*}
\interpApprox{-} & : \textit{approx} \rightarrow (A \rightarrow
  \textsf{Approx} \, A) \\
\interpApprox{\texttt{atLeast}} & = \mathsf{lower} \\
  \interpApprox{\texttt{atMost}} & = \mathsf{upper} \\
  \interpApprox{\epsilon} & = \mathsf{exact} \\ \\
  \interpMult{-} & : \textit{mult} \rightarrow (A \rightarrow
  \textsf{Mult} \, A) \\
  \interpMult{\texttt{readOnce}} & = \mathsf{once} \\
  \interpMult{\epsilon} & = \mathsf{mult}
\end{align*}
\caption{Interpretation of modifiers}
\label{subfig:modifier-model}
\end{subfigure}
\caption{Semantic model of specifications}
\end{figure}

The intermediate interpretation $\interp{-}^{\mathsf{c}}$ of
\Cref{subfig:region-model} models region contants. This is lifted by
$\textit{promote}_N$ to interpret region constants in the last
equation of $\interp{-}_N$. The \term{+} and \term{*} operators are
modelled in terms of the join (union) and meet (intersection) of
interval schemes. Thus the syntax of regions is modelled as members of
$\region{N}$.

We mark in our model the presence of modifiers such as
\texttt{readOnce} and \texttt{atMost} as introduced in
\Cref{defn:modifiers}.  Approximation modifiers are interpreted as
injections into the $\mathsf{Approx}$ variant by $\interpApprox{}$ in
\Cref{subfig:modifier-model}.  The $\textsf{Approx}$ type corresponds
to the presence or absence of the spatial approximation modifier, with
\textsf{exact} when there is no such modifier and \textsf{lower} and
\textsf{upper} for \term{atLeast} and \term{atMost}. In a similar way,
multiplicity modifiers are interpreted as injections in the
$\mathsf{Mult}$ variant by $\interpMult{}$, corresponding to the to
the presence or absence of the \term{readOnce} modifier as shown in
\Cref{subfig:modifier-model}.

\begin{defn}[Semantics of specifications]
The intermediate interpretations of \Cref{fig:semantics}
are composed to give a model for the top-level specification
syntax as:
%
\begin{equation*}
  \interp{\texttt{stencil \textcap{mult}, \textcap{approx}, \textcap{region}}}_N =
    \interpMult{\textcap{mult}} \;
           {(\interpApprox{\textcap{approx}} \;
                    {\interp{\textcap{region}}_N)}}
\end{equation*}
\end{defn}

\begin{restatable}[Equational soundness]{thm}{equationalSoundness}
  The lattice model is sound with respect to the equational theory. Let $R$ and $S$
  be $N$-dimensional region terms, then we have
%
  \begin{equation*}
    \forall R, S, N. \; R \equiv S \implies \interp{R}_N = \interp{S}_N
  \end{equation*}
\end{restatable}

\begin{restatable}[Approximation soundness]{thm}{approxSoundness}
  The lattice model is sound with respect to the theory of approximation. Let $R$
  and $S$ be $N$ dimensional regions, then we have
%
  \begin{equation*}
    \forall R, S, N. \; R \preceq S \implies \interp{R}_N \subseteq \interp{S}_N
  \end{equation*}
\end{restatable}
%
Note that the model is not complete with respect to equations or
approximations since the specification language has no model of the
bottom element of the lattice.

\subsection{Interpreting array subscripts}
\label{sec:fromcode}

\begin{defn}
  \label{defn:subscript}
  Recall array subscript terms of the form $a(\bar{e})$ from
  \Cref{def:array-subs}. We interpret these terms with the partial
  interpretation $\interp{-}^{\mathit{aterm}} : \textit{array-term}
  \prightarrow{} \mathcal{P}(\bz{})^N$. The interpretation is defined when
  all indices are either constant or neighbourhood indices as defined in
  \Cref{def:neighbour-ix}.
%
  \begin{align*}
    \interp{a(\bar{e})}^{\mathit{aterm}} =
      \prod_{1 \leq i \leq N} \mathit{subscript}(\bar{e}_i) & &
  %
    \textit{subscript}(e) = \begin{cases}
      \{ c \} & e \equiv i \pm c \\
      \bz & e \; \mbox{is absolute}
    \end{cases}
  \end{align*}
\end{defn}
%
Note that this produce subscript schemes (a form of index scheme), but
they are not members of \textit{Interval} or $\textit{Region}_N$. In
the next section, we use this interpretation on array subscripts in
the static analysis producer preceding checking and inference.

\subsection{Union Normal Form}\label{subsec:union-normal-form}

\noindent
We lastly address a problem of \emph{representation} for index schemes
in the model. Index schemes may incorporate the infinite set \bz{} in
their components, \eg{} when a dimension is unconstrained or when an
absolute index of array term is interpreted. In order to provide a
finite checking procedure, we require a finite and compact
representation for indices which accounts for this.

Recall \Cref{lem:vector-intersect} states that the
intersection of indexing schemes is still an indexing scheme. When
combined with \Cref{lem:interval-identities} (union and intersection
on holed intervals), we can extend this closure to interval schemes.
Subscripts schemes do not enjoy the same propery as intersection of
two singleton sets might easily be empty, which disqualifies it from
being a subscript index. In the following sections we will not take
intersection of subscript indices.

Union, however, is not as flexible as
intersection. \Cref{lem:vector-union} does not form a closure for
indexing schemes in general. For this reason, we represent sets of
indexing schemes as unevaluated union terms, or \emph{union normal
  form}.  Since the model forms a distributive lattice as established
in \Cref{prop:region-lattice}, the intersections can be pushed inwards
using the distributive law leaving only unions at the outer-level. We
further exploit associativity of union to put unions into a cons-tree
making the unions effectively a non-empty list.  We do not, however,
attempt to canonicalise the representation. Later on
\Cref{lem:vector-union} and index scheme subset are used to reduce the
size of union normal form.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: