### 2017-04-18 oopsla17/conclusions.tex
various edits, shrinking, and conclusion

\textbf{Concluding remarks} \hspace{0.25em}
We proposed a novel solution to a common programming task in
numerical and scientific computing. Our approach to language design is
also relatively uncommon-- informed by a quantitative study of
existing code. The resulting language is flexible and expressive,
capturing roughly 80,000 array computations across a million lines of
Fortran with simple, short, abstract specifications. %As
%opposed to existing approaches our language takes inspiration from the
%numerical literature, is derived from a quantitative analysis of
%real-world acode, and provides a concise, abstract description of the
%stencil.
We are now in the
process of applying our tool in collaboration with
the authors' of some of our corpus packages.
Future work is to expand the range of our program analysis, including
options for further semantic analysis and capturing a wider set of
patterns.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: